# Setup

## Load libraries

```{r load_libs}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(here)           # To work with paths
```

# Read data

## Read annex data

### GBIF matched annex data in BIM list

Import `full_match_annexes.csv`:

```{r read_annexes}
annex_in_bim <- read_delim(
  file = here("data", "interim", "full_match_annexes.csv"), 
  delim = ",")
```

### GBIF matched annex data not in BIM list

Import `taxa_annex_not_in_bim` 

```{r read_annexes}
annex_not_in_bim <- read_delim(
  file = here("data", "interim", "taxa_annex_not_in_bim.csv"), 
  delim = ",")
```

### GBIF unmatched annex data

```{r}
annex_unmatched_taxa <- read_csv(
  here("data", "processed", "unmatched_taxa_annexes.csv"),
  na = "",
  col_types = cols(
    .default = col_character()
))
```

### Import BIM annex table:

```{r read_taxa_parentid_corrected}
taxa_parentid_corrected <- read_delim(
  file = here("data", "processed", "taxa_parentid_corrected.tsv"), 
  delim = "\t",
  col_types = cols(
    .default = col_character(),
    id = col_double(),
    parentid = col_double(),
    media = col_logical(),
    speciesbeparentid = col_double(),
    bruenvi_created = col_date(format = "%Y-%m-%d"),
    bruenvi_modified = col_date(format = "%Y-%m-%d"))
)
```

# Generate revised speciesannex table

Process `unmatched_bim_annex_taxa` for future join with `annex_in_bim` and `annex_not_in_bim`:

1. Add column `gbif_scientificName` which is empty for unmatched taxa

```{r}
annex_unmatched_taxa <-
  annex_unmatched_taxa %>% 
    select(-scientific_name_corrected) %>% 
    mutate(gbif_scientificName = "")
```

2. Add column  `relation_to_annexspecies`:

```{r}
annex_unmatched_taxa <-
  annex_unmatched_taxa %>% 
    mutate(relation_to_annexspecies = "not in GBIF backbone")
```

Bind rows to generate annexspecies_new:

```{r}
annexspecies_new <-
  bind_rows(annex_in_bim, 
            annex_not_in_bim,
            annex_unmatched_taxa)
```

Add `taxonid` from revised BIM taxon table:

```{r}
annexspecies_new <- 
  annexspecies_new %>% 
    left_join(
      select(taxa_parentid_corrected, id, gbif_scientificName),
      by = c("gbif_scientificName"))
```

Restructure `annexspecies_new`:

```{r}
annexspecies_new <- 
  annexspecies_new %>% 
    select(id,
           annex_code,
           scientific_name_original,
           gbif_scientificName,
           relation_to_annexspecies)
```

Recode `annex_code` to codes used in BIM annex list:

```{r}
annexspecies_new <- 
  annexspecies_new %>% mutate(annex_code = recode(annex_code,
      "bxl_2.1" = "BXL-ORD-2012_Annex II.1",
      "bxl_2.2" = "BXL-ORD-2012_Annex II.2",
      "bxl_2.3" = "BXL-ORD-2012_Annex II.3",
      "bxl_2.4" = "BXL-ORD-2012_Annex II.4",
      "bxl_2.5" = "BXL-ORD-2012_Annex II.5",
      "bxl_3" = "BXL-ORD-2012_Annex III",
      "bxl_4" = "BXL-ORD-2012_Annex IV",
      "bern_1" = "EUR-CON-BER_Annex I",
      "bern_2" = "EUR-CON-BER_Annex II",
      "bern_3" = "EUR-CON-BER_Annex III",
      "bonn_1" = "UN-CON-BON-Annex I",
      "bonn_2" = "UN-CON-BON-Annex II",
      "hab_2" = "EUR-DIR-HAB_Annex II",
      "hab_4" = "EUR-DIR-HAB_Annex IV",
      "bird_1" = "EUR-DIR-BIR_Annex I"))
```

# Export speciesannex_new

```{r}
annexspecies_new %>% 
  write_csv(path = here("data", "processed", "speciesannex_new.csv"),
            na = "")
```

