---
title: "Add taxonomic information to BIM annex data"
author:
  - Lien Reyserhove
  - Damiano Oldoni
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Setup

## Load libraries

```{r load_libs}
library(odbc)           # To work with database
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(here)           # To work with paths
library(rgbif)          # To work with GBIF data
```

## Load access informations

Retrieve access informations from configuration file:

```{r get_access_infos}
ibge_bim <- config::get("ibge_bim")
```

Connect to database:

```{r connect_to_db}
conn <- dbConnect(odbc(), 
                  driver = ibge_bim$driver,
                  server = ibge_bim$server,
                  database = ibge_bim$database,
                  port = ibge_bim$port,
                  uid = ibge_bim$uid,
                  pwd = ibge_bim$pwd,
                  encoding = "Windows-1252"
)
```

# Read data

We define first the annex codes we will use:

```{r annex_codes}
annexcode <- c("bxl_2.1",
               "bxl_2.2",
               "bxl_2.3",
               "bxl_2.4",
               "bxl_2.5",
               "bxl_3",
               "bxl_4",
               "bern_1",
               "bern_2",
               "bern_3",
               "bonn_1",
               "bonn_2",
               "hab_2",
               "hab_4",
               "bird_1")
```

## BIM annex data

Extract annex information from the database (`biodiv.speciesannex`):

```{r}
bim_annex_species <- dbGetQuery(conn, "
  SELECT
    sa.taxonid,
    sa.annexcode,
    a.descriptionnl
  FROM biodiv.speciesannex AS sa
    LEFT JOIN biodiv.annex AS a
      ON sa.annexcode = a.annexcode
  ORDER BY
    sa.annexcode,
    sa.taxonid
") %>% 
  as_tibble()
```

Overview of the BIM annexes:

```{r overview_bim_annex_species}
bim_annex_species %>% distinct(annexcode) 
```

## BIM taxonomic data

Import `taxa_parentid_corrected.tsv` to link `taxonid` with scientific names and related gbif taxonomic information:

```{r read_taxa_parentid_corrected}
taxa_parentid_corrected <- read_delim(
  file = here("data", "processed", "taxa_parentid_corrected.tsv"), 
  delim = "\t",
  col_types = cols(
    .default = col_character(),
    id = col_double(),
    parentid = col_double(),
    media = col_logical(),
    speciesbeparentid = col_double(),
    bruenvi_created = col_date(format = "%Y-%m-%d"),
    bruenvi_modified = col_date(format = "%Y-%m-%d"))
)
```

# Add taxonomic information to BIM annex

## Join biodiv.taxon with BIM annex

Join the taxonomic information from `taxa_parentid_corrected` with `bim_annex_species`:

```{r join_bim_annex_species_taxa_parentid_corrected}
bim_annex_species  %<>% 
  left_join(taxa_parentid_corrected, by = c("taxonid" = "id")) %>% 
  arrange(taxonid)
```

Inspect `gbif_matchType` for nonmatching taxa:

```{r inspect_gbif_matchType}
bim_annex_species %>% 
  group_by(gbif_matchType) %>% 
  count()
```

## Get names of accepted taxa from GBIF

In BIM annexes there are some synonyms, based on GBIF Taxonomy Backbone:

```{r n_synonyms}
bim_annex_species %>%
  filter(gbif_matchType != "NONE") %>%
  group_by(gbif_synonym) %>%
  count()
```

Find accepted (scientific) names of the taxa this synonyms point to and add them in new column `gbif_accepted_name`. For accepted taxa, we set `gbif_accepted_name` equal to `gbif_scientificname`:

```{r get_accepted_names_bim_annexes}
bim_annex_species <- 
  bim_annex_species %>%
  mutate(gbif_accepted_name = map2_chr(
    bim_annex_species$gbif_acceptedUsageKey,
    bim_annex_species$gbif_scientificName,
    function(key, gbif_scientificname){
      if (!is.na(key)) {
        name_usage(key, return = "data")$scientificName
      } else {
        gbif_scientificname
      }
    })
)
```

## Save BIM annex data

We save `bim_annex_species` in  `/data/interim`:

```{r save-bim_annex_species}
write_csv(
  bim_annex_species,
  path = here("data", "interim", "bim_annex_taxa_full_taxonomic_info.csv"),
  na = ""
)
```
