---
title: "Taxa: retrieve kingdom IDs and detect anomalies"
author:
  - Damiano Oldoni
  - Lien Reyserhove
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

# Setup

## Load libraries

```{r add_kingdom-load_libs}
library(tidyverse)  # To do data science
library(tidylog)    # To provide feedback on dplyr functions
library(here)       # To work with paths
```

# Read taxa

Read taxa from `./data/input/taxa.tsv`:

```{r add_kingdom-read_taxa_from_tsv_file}
taxa <- read_tsv(here("data", "input", "taxa.tsv"),
                 col_types = cols(
                   .default = col_character(),
                   id = col_double(),
                   parentid = col_double(),
                   media = col_logical(),
                   speciesbeparentid = col_double(),
                   bruenvi_created = col_date(format = "%Y-%m-%d"),
                   bruenvi_created = col_date(format = "%Y-%m-%d"))
)
```

# Overview ranks and parent IDs

Taxonomy ranks present in `taxa`:

```{r add_kingdom-get_taxonranks}
taxa %>% 
  group_by(taxonranken) %>% 
  summarize(records = n())
```

Taxa without parent ID:

```{r add_kingdom-n_taxa_with_no_parentid}
taxa %>%
  group_by(!is.na(parentid)) %>%
  summarize(records = n()) %>%
  rename( has_parent = "!is.na(parentid)")
```


Check the internal validity of parent IDs. Are all parent IDs valid IDs?

```{r add_kingdom-check_parent_IDs}
parent_ids <- taxa %>%
  filter(!is.na(parentid)) %>%
  pull(parentid)
all(parent_ids %in% taxa$id)
```

Taxon ranks linked to the parent IDs:

```{r add_kingdom-taxon_ranks_of_parents}
taxa %>%
  filter(id %in% parent_ids) %>%
  group_by(taxonranken) %>% 
  summarize(records = n())
```

This means that the following taxon ranks are not pointed out by any taxa:

```{r add_kingdom-not_pointed_out_ranks}
taxonranks <- 
  taxa %>% 
  distinct(taxonranken)
not_pointed_out_taxonranks <- 
  taxonranks %>%
  anti_join(
    taxa %>% 
      filter(id %in% parent_ids) %>%
      distinct(taxonranken),
    by = "taxonranken"
)
not_pointed_out_taxonranks
```

This reflects the structure of GBIF Taxonomy backbone (see issue https://github.com/gbif/checklistbank/issues/79) and Catalogue of Life, where no `subfamily`, `superfamily`, `superorder`, `suborder` or `informal group` are present in the upper hierarchy. On the other hand, the fact that `forma` and `subforma` are never parents of other taxa is quite reasonable. In the case of `subforma`, there refer to `species` (and thus not `forma`) as a taxon rank which is ok.  

# Search taxa in GBIF Taxonomy Backbone

We will search each taxa within the GBIF Taxonomy Backbone. It is a search *by scientific name*. However, homonyms are present among different kingdoms. This means that a scientific name can occur in different kingdoms, these are so-called hemihomonyms. For example the name _Erica_ has been given to both a genus of spiders, `Erica Peckham & Peckham, 1892` and to a genus of heaths `Erica L.`. To avoid a taxon being misidentified, we first need to retrieve the kingdom of each taxon. 

## Get kingdom 

We write a recursive function, called `get_keys_higher_levels`, to keep track of all keys while *climbing up* to the highest possible level for each taxon. This function uses the parent IDs until no parent ID is found anymore, meaning we arrived at kingdom taxonomic level if no mising parent keys are found.  The function returns a vector containing the sequence of parent IDs. If a taxa is found where `parentid` = `id` or the sequence of parent IDs is circular, the vector contains `NA` at first position in order to be easily found for further inspection.

```{r add_kingdom-function_get_kingdom}
#' Function to retrieve all parent taxa up to kingdom recursively via parent ID's.
#' 
#' @param taxon_id (double) A unique ID identifying taxon in `taxa_df`.
#' @param taxa_df (tibble) A data.frame containing the taxonomic data. It MUST
#' contains the following columns: `id` and `parent_id`.
#' @param parents_vector (double) A vector of parent IDs. It is populated within recursion. Default: NULL.
#' @return A vector of parent IDs. If a corrupted parent IDs is found, the
#' vector contains NA at first position.
get_keys_higher_levels <- function(taxon_id, taxa_df = taxa, parents_vector = NULL) {
  parent_id <- taxa_df %>%
    filter(id == taxon_id) %>%
    pull(parentid)
  parents_vector <- c(taxon_id, parents_vector)
  if (parent_id %in% parents_vector) {
      return(c(NA, parents_vector))
  } else {
    if (is.na(parent_id)) {
      return(parents_vector)
    } else {
      return(c(get_keys_higher_levels(taxon_id = parent_id,
                                        taxa_df =  taxa_df,
                                        parents_vector = parents_vector)))
    }
  }
}
```

We first initialize a new empty column called `kingdom_id` in `taxa`:

```{r add_kingdom-initialize_col_kingdom_id_taxa}
taxa <- 
  taxa %>%
  mutate(kingdom_id = 0)
```

We apply the function `get_keys_higher_levels` to all taxa. If the highest taxonomic level pointed by the sequence of parent IDs is not equal to `kingdom` or the vector of parent IDs starts with `NA`, `kingdom_id` will be set to `NA`:

```{r add_kingdom-find_kingdom_id_all_taxa, cache=TRUE}
circular_parent_ids <- c()
pb <- txtProgressBar(min = 0, max = nrow(taxa),  style = 3)
for (t in 1:nrow(taxa)) {
  if (!is.na(taxa$kingdom_id[t]) & taxa$kingdom_id[t] == 0) {
    higher_levels_ids <- get_keys_higher_levels(taxon_id = taxa$id[t], taxa, NULL)
    if (!is.na(higher_levels_ids[1])) {
      taxon_rank_highest_level <-
        taxa$taxonranken[taxa$id == higher_levels_ids[1]]
    } else {
      circular_parent_ids <- c(list(higher_levels_ids[2:length(higher_levels_ids)]),
                               circular_parent_ids)
      for (i in higher_levels_ids[2:length(higher_levels_ids)]) {
        taxa$kingdom_id[taxa$id == i] <- NA_real_
      }
    }
    for (i in higher_levels_ids) {
      if (taxon_rank_highest_level == "kingdom") {
        taxa$kingdom_id[taxa$id == i] <- higher_levels_ids[1]
      } else {
        taxa$kingdom_id[taxa$id == i] <- NA_real_
      }
    }
  }
  setTxtProgressBar(pb, t)
}
```

Kingdoms present in `taxa`:

```{r add_kingdom-add_kingdom_names}
kingdom_df <- 
  taxa %>%
  filter(taxonranken == "kingdom") %>%
  select(id, acceptedname) %>%
  rename(kingdom = acceptedname)
kingdom_df
```

We add `kingdom` to `taxa` for better interpretation of the results. We call this column `bim_kingdom` in order to distinct it in the future from the `kingdom` got by GBIF Taxonomy Backbone:

```{r add_kingdom-add_kingdom_column}
taxa <- 
  taxa %>%
  left_join(kingdom_df, by = c("kingdom_id" = "id")) %>%
  rename(bim_kingdom = kingdom)
```

Preview of taxa linked to their specific kingdom:

```{r add_kingdom-preview_taxa_with_kingdom}
taxa %>% 
  select(acceptedname, bim_kingdom) %>% 
  head(n = 100)
```

Number of taxa without `kingdom_id`:

```{r add_kingdom-invalid_kingdom_id}
taxa %>%
  filter(is.na(kingdom_id)) %>%
  nrow()
```

Some of these ones have parent IDs ending up in a circular sequence. Blindly following the path of parent IDs would then result in an infinite loop:

```{r add_kingdom-circular_parent_ids}
tibble(id = unlist(circular_parent_ids)) %>%
  left_join(taxa, by = "id") %>%
  select(acceptedname, id, parentid)
```

In our case the *Salmo trutta fario* refers to itself, `parentid` =  `id`, so anything linking directly or indirectly to this taxon will not get any step further.

We correct the parent ID of *Salmo trutta fario* immediately:

```{r add_kingdom-correct_circular_parent_ids}
taxa <- 
  taxa %>%
  mutate(parentid = case_when(
    acceptedname == "Salmo trutta fario" ~ 47683,
    TRUE ~ parentid)
)
```

# Save taxa

We save all taxa in `.\data\interim`:

```{r add_kingdom-save_all_taxa}
write_tsv(taxa, 
          here("data", "interim", "taxa.tsv"), 
          na = "")
```
