---
title: "Detect unused taxa"
author:
  - Damiano Oldoni
  - Lien Reyserhove
  - Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

# Setup

## Load libraries

```{r load_libs}
library(odbc)       # To work with database
library(tidyverse)  # To do data science
library(here)       # To work with paths
library(inborutils) # To match taxa (rgbif called in background)
```

## Load access informations

Retrieve access informations from configuration file:

```{r get_access_infos}
ibge_bim <- config::get("ibge_bim")
```

## Connect to database

Connect to database:

## Connect to database

Connect to database:

```{r connect_to_db}
conn <- dbConnect(odbc(), 
                  driver = ibge_bim$driver,
                  server = ibge_bim$server,
                  database = ibge_bim$database,
                  port = ibge_bim$port,
                  uid = ibge_bim$uid,
                  pwd = ibge_bim$pwd,
                  encoding = "Windows-1252"
)
```

# Read data

We read first taxonomic information from file `taxa` containing a copy of table `taxon`. Next, we retrieve information from the tables which are linked to these taxa:

1. table `identifiablespecies`
2. table `speciesannex`
3. table `media`
4. table `taxoncommontaxa`

We also retrieve occurrences from table `occurence` in order to assess which taxa are not linked to any occurrence.

## Read taxonomic data

We read taxa. They were previously saved as text file:

```{r import_taxa_from tsv}
taxa <- read_tsv(here("data", "input", "taxa.tsv"),
                 col_types = cols(
                   .default = col_character(),
                   id = col_double(),
                   parentid = col_double(),
                   media = col_logical(),
                   speciesbeparentid = col_double(),
                   bruenvi_created = col_date(format = "%Y-%m-%d"),
                   bruenvi_modified = col_date(format = "%Y-%m-%d"))
)
```

## Read occurrence related data

### Read occurrences

Read table `occurence` containing occurrence data:

```{r read_occurence_table, cache=TRUE}
occurrences <- 
  dbGetQuery(conn, "SELECT * FROM biodiv.occurence") %>%
  as_tibble()
```

Preview:

```{r occurrences_preview}
head(occurrences)
```

### Read intermediate table `identifiablespecies`

The taxonomic information of occurrences is linked to taxa via intermediate table `identifiablespecies` where a many-to-one relation occurs. This is the way synonym relations are handled. Read data from `identifiablespecies`:

```{r identifiable_species, cache=TRUE}
identifiable_species <- 
  dbGetQuery(conn, "SELECT * FROM biodiv.identifiablespecies") %>%
  as_tibble()
```

Preview:

```{r preview_identifiablespecies}
head(identifiable_species)
```

The `id` column contains values present in `occurrences$identifiablespeciesid`, while `taxonid` contains values present in `taxa$id`.

Notice the many-to-one relation. Example: `taxonid` 79:

```{r example_taxonid_79}
identifiable_species %>%
  filter(taxonid == 79)
```

## Read annex data

```{r get_annex_data, cache=TRUE}
annex <- 
  dbGetQuery(conn, "SELECT * FROM biodiv.speciesannex") %>%
  as_tibble()
```

Preview:

```{r preview_speciesannex}
head(annex)
```

Taxonomic information is saved in column `taxonid`.

## Read media data

Read media from table `media`:

```{r get_media, cache=TRUE} 
media <- 
  dbGetQuery(conn, "SELECT * FROM biodiv.media") %>%
  as_tibble()
```

Preview:

```{r preview_media}
head(media)
```

## Read common taxa

Read common taxa from table `taxoncommontaxa`:

```{r get_taxoncommontaxa, cache=TRUE}
taxon_common_taxa <- 
  dbGetQuery(conn, "SELECT * FROM biodiv.taxoncommontaxa") %>%
  as_tibble()
```

Preview:

```{r preview_taxon_common_taxa}
head(taxon_common_taxa)
```

As in previous tables, taxonomic information is saved in column `taxonid`. The column  `commontaxaid` contains the identifier to define the typic taxonoic groups to search on. This information is saved in table `commontaxa`:

```{r get_commontaxa, cache=TRUE}
common_taxa <- 
  dbGetQuery(conn, "SELECT * FROM biodiv.commontaxa") %>%
  as_tibble()
```

Preview:

```{r preview_common_taxa}
head(common_taxa)
```

# Detect (un)used taxa

## Taxa not in occurrences

Taxa not linked to any occurrence are found by using information in intermediate table `identifiablespecies`:

```{r taxa_not_in_occs}
taxa_not_in_occs <- 
  taxa %>%
  anti_join(
    occurrences %>%
    distinct(identifiablespeciesid) %>%
    left_join(identifiable_species, 
              by = c("identifiablespeciesid" = "id")) %>%
    select(taxonid),
    by = c("id" = "taxonid")) %>%
  select(names(taxa))
```

Number of taxa linked to occurrences:

```{r n_taxa_linked_to_occs}
taxa %>%
  mutate(used_for_occs = if_else(id %in% taxa_not_in_occs$id,
                                 FALSE, TRUE)) %>%
  group_by(used_for_occs) %>%
  summarize(n_taxa_with_occs = n())
```

## Taxa not used in annexes

Taxa not included in any annex:

```{r taxa_not_in_annex}
taxa_not_in_annex <-
  taxa %>%
  anti_join(annex %>% distinct(taxonid),
            by = c("id" = "taxonid"))
```

Number of taxa linked to annexes:

```{r n_taxa_linked_to_annexes}
taxa %>%
  mutate(used_for_annex = if_else(id %in% taxa_not_in_annex$id,
                                 FALSE, TRUE)) %>%
  group_by(used_for_annex) %>%
  summarize(n_taxa_in_annex = n())
```

## Taxa not used in media

Taxa not linked to media:

```{r taxa_not_in_media}
taxa_not_in_media <-
  taxa %>%
  anti_join(media %>% distinct(taxonid),
            by = c("id" = "taxonid"))
```

Number of taxa with media:

```{r n_taxa_with_media}
taxa %>%
  mutate(has_media = if_else(id %in% taxa_not_in_media$id,
                                 FALSE, TRUE)) %>%
  group_by(has_media) %>%
  summarize(n_taxa_with_media = n())
```

## Taxa not in common groups

Taxa not belonging to any common group:

```{r taxa_not_in_common_groups}
taxa_not_in_common_groups <- 
  taxa %>%
  anti_join(taxon_common_taxa %>% distinct(taxonid),
            by = c("id" = "taxonid"))
```

Number of taxa not in common groups:

```{r n_taxa_belong_to_groups}
taxa %>%
  mutate(in_groups = if_else(id %in% taxa_not_in_common_groups$id,
                                 FALSE, TRUE)) %>%
  group_by(in_groups) %>%
  summarize(n_taxa_in_groups = n())
```

## Used and unused taxa

By inner join on previous results we can find which taxa are not used neither in occurrences nor annex nor common groups nor media:

```{r not_used_taxa}
not_used_taxa <- 
  taxa_not_in_occs %>% 
  filter(id %in% taxa_not_in_annex$id) %>%
  filter(id %in% taxa_not_in_common_groups$id) %>%
  filter(id %in% taxa_not_in_media$id)
```

Complementarily, taxa used at least by one of the linked tables:

```{r used_taxa}
used_taxa <- 
  taxa %>%
  filter(!id %in% not_used_taxa$id)
```

Parents of the used taxa:

```{r parents_used_taxa}
parents_used_taxa <-
  taxa %>%
  filter(id %in% 
           (used_taxa %>% distinct(parentid) %>% pull()))
```

We remove these taxa from the unused taxa:

```{r not_unused_taxa_not_parents}
not_used_taxa <-
  not_used_taxa %>%
  filter(!id %in% parents_used_taxa$id)
```

Final number of not used taxa:

```{r nrow_not_used_taxa}
nrow(not_used_taxa)
```

# Save data

## Save occurrence data

Save occurrences locally, but not push to repo via `.gitignore`:

```{r save_occurrences}
write_tsv(occurrences,
          path = here("data", "input", "occurrences.tsv"),
          na = "")
```

## Save identifiable species

```{r save_identifiable_species}
write_tsv(identifiable_species,
          path = here("data", "input", "identifiable_species.tsv"),
          na = ""
)
```

## Save taxa

### Used taxa

We save also the used taxa as these ones will be the nucleus of the new taxa table:

```{r save_used_taxa}
write_tsv(used_taxa,
          path = here("data", "interim", "used_taxa.tsv"),
          na = "")
```

### Not used taxa

Save taxa which are not currently in use:

```{r save_not_used_taxa}
write_tsv(not_used_taxa,
          path = here("data", "interim", "not_used_taxa.tsv"),
          na = "")
```
