---
title: "Final check of taxon table and assign date creation"
author:
  - Lien Reyserhove
  - Damiano Oldoni
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Setup

## Load libraries

```{r load_libs}
library(odbc)           # To work with database
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(here)           # To work with paths
library(rgbif)          # To work with GBIF data
```

# Read data

We read taxonomic information by importing the  `biodiv.taxon.txt` table and its extended version, `taxa_parentid_corrected_unverifiable_taxa_added.tsv`.

```{r read_biodiv_taxon}
biodiv_taxon <- read_delim(here("data", "processed", "biodiv.taxon.txt"),
                           delim = "|",
                           na = "",
                           col_types = cols(
                             .default = col_character(),
                             id = col_double(),
                             parentid = col_double()
                           ))
```

```{r biodiv_taxon_extended}
biodiv_taxon_extended <- read_tsv(
  here("data", 
       "processed", 
       "taxa_parentid_corrected_unverifiable_taxa_added.tsv"),
  na = "",
  col_types = cols(
    .default = col_character(),
    id = col_double(),
    parentid = col_double(),
    media = col_logical(),
    speciesbeparentid = col_double(),
    bruenvi_created = col_date(format = "%Y-%m-%d"),
    bruenvi_modified = col_date(format = "%Y-%m-%d"),
    gbif_confidence = col_double(),
    gbif_usageKey = col_double(),
    gbif_speciesKey = col_double(),
    gbif_genusKey = col_double(),
    gbif_parentKey = col_double(),
    gbif_familyKey = col_double(),
    gbif_classKey = col_double(),
    gbif_orderKey = col_double(),
    gbif_phylumKey = col_double(),
    gbif_kingdomKey = col_double(),
    gbif_acceptedUsageKey = col_double(),
    gbif_synonym = col_logical(),
    added_from_gbif = col_logical())
)
```

# Solve duplicates

## Get duplicates in  `biodiv.taxon` 

Duplicates based on `acceptedname`:

```{r duplicate_names}
duplicate_names <- 
  biodiv_taxon %>%
  group_by(acceptedname, scientificnameauthorship) %>%
  count() %>%
  filter(n > 1) %>%
  ungroup()
```

Show duplicates:

```{r show_duplicate_names}
duplicate_names <- 
  duplicate_names %>% 
  left_join(biodiv_taxon, 
             by = c("acceptedname", "scientificnameauthorship")) %>%
  select(id, acceptedname, scientificnameauthorship, parentid, everything())
duplicate_names
```

The presence of duplicates is already known and documented in [#7](https://github.com/inbo/ibge-bim-species/issues/7). 

## Find used taxa

We are going to remove unused duplicates only, so without changing anything in linekd tables.

### Get IDs used in linked tables

Based on [biodiv_diagram](https://github.com/inbo/ibge-bim-species/blob/master/references/biodiv_diagram.pdf) the following tables of `biodiv` are linked to `biodiv.taxon`:

1. `identifiablespecies`
2. `media`
3. `speciesannex`
4. `taxoncommontaxa`

Differently from pipeline `parentid_2_detect_used_taxa.Rmd`, we are less restricted in asssing the usage of a taxon: we don't check which names in `identifiablespecies` are really linked to occurrences. Again, this is done to avoid to remove names in `biodiv.taxon` which are used  `identifiablespecies`.

Retrieve access informations from configuration file:

```{r get_access_infos}
ibge_bim <- config::get("ibge_bim")
```

Connect to database:

```{r connect_to_db}
conn <- dbConnect(odbc(), 
                  driver = ibge_bim$driver,
                  server = ibge_bim$server,
                  database = ibge_bim$database,
                  port = ibge_bim$port,
                  uid = ibge_bim$uid,
                  pwd = ibge_bim$pwd,
                  encoding = "Windows-1252"
)
```

Get all taxonIDs used in `identifiablespecies`:

```{r get_ids_identifiablespecies}
taxonid_identifiablespecies <- 
  dbGetQuery(conn, 
             "SELECT DISTINCT taxonid FROM biodiv.identifiablespecies") %>%
  as_tibble()
```

Number of unique taxon IDs:

```{r n_taxa_identifiablespecies}
nrow(taxonid_identifiablespecies)
```

Get all taxonIDs used in `media`:

```{r get_ids_identifiablespecies}
taxonid_media <- 
  dbGetQuery(conn, 
             "SELECT DISTINCT taxonid FROM biodiv.media") %>%
  as_tibble()
```

Number of unique taxon IDs:

```{r n_taxa_media}
nrow(taxonid_media)
```

Get all taxonIDs used in `speciesannex`:

```{r get_ids_speciesannex}
taxonid_speciesannex <- 
  dbGetQuery(conn, 
             "SELECT DISTINCT taxonid FROM biodiv.speciesannex") %>%
  as_tibble()
```

Number of unique taxon IDs:

```{r n_taxa_speciesannex}
nrow(taxonid_speciesannex)
```

Get all taxonIDs used in `taxoncommontaxa`:

```{r get_ids_taxoncommontaxa}
taxonid_taxoncommontaxa <- 
  dbGetQuery(conn, 
             "SELECT DISTINCT taxonid FROM biodiv.taxoncommontaxa") %>%
  as_tibble()
```

Number of unique taxon IDs:

```{r n_taxa_taxoncommontaxa}
nrow(taxonid_taxoncommontaxa)
```

Close connection:

```{r close_connection}
dbDisconnect(conn)
```

We also get taxon IDs used as parent IDs in `biodiv.taxon`. But for these IDs we use our final `biodiv.taxon.txt`.

```{r get_parentids_taxon}
taxon_parentid <- 
  biodiv_taxon %>% 
  distinct(parentid)
```

Get all unique taxon IDs from all tables as a vector, `taxonids`: 

```{r all_used_taxonids}
taxonids <- c(taxonid_identifiablespecies %>% pull(taxonid),
              taxonid_media %>% pull(taxonid),
              taxonid_speciesannex %>% pull(taxonid),
              taxonid_taxoncommontaxa %>% pull(taxonid),
              taxon_parentid %>% pull(parentid))
taxonids <- unique(taxonids)
```

## Unused duplicates

Now we can remove any taxa in `biodiv.taxon` which is not in `taxonids` as they are never used.

We show here below the unused duplicates, labelled  as `used` =  `FALSE` together with the used duplicates, labelled as `used` =  `TRUE`

```{r not_used_duplicates}
not_used_duplicate_names <-
  duplicate_names %>%
  filter(!id %in% taxonids & taxonranken != "kingdom") %>%
  mutate(used = FALSE)
not_used_duplicate_names
used_duplicate_names <-
  duplicate_names %>%
  filter(id %in% taxonids & 
           acceptedname %in% not_used_duplicate_names$acceptedname &
           scientificnameauthorship  %in% not_used_duplicate_names$scientificnameauthorship) %>%
  mutate(used = TRUE)
bind_rows(not_used_duplicate_names,
          used_duplicate_names) %>%
  arrange(acceptedname, scientificnameauthorship) %>%
  select(id, 
         acceptedname, 
         scientificnameauthorship, 
         parentid, 
         used, 
         everything())
```

We discuss the different patterns we can see and we explain the reason why these duplicates arise.

### Duplicates already present in biodiv.taxon

Used and unused taxa already present in original version of  `biodiv.taxon`, i.e. `id` less than 100000. We speak about the following taxa:

1. Botrytella, 
2. Dasysiphonia, 
3. Macrosiagon, 
4. Myriactula, 
5. Saccharomyces bayanus, 
6. Saccharomyces cerevisiae, 
7. Schizophyllum

### Carlia, Chondromyces, Echiuroidea

These duplicates share following properties:

1. unused original taxon
2. no parent ID in unused original taxon
2. no match to GBIF Backbone (multiple equal matches)
3. used taxon added by building txonomic tree
4. used taxon without authorship

### Characium, Dicranochaete, Euzonus, Grandinia, Thorea

These duplicates have the following properties:
1. original taxon not used
2. added taxon used (`id` higher than 100000)
3. parent ID corrected
4. genus or higher rank

This happens due to a successful match with GBIF Backbone which reveals afterhand to be not the right one. As a consequence, their parent ID has been corrected to a wrong one and their children are therefore redirected to the new created taxon, thus avoiding to introduce any error in  `biodiv.taxon`. They are not used as parents, while their added counterparts are well used.

```{r errors_in_gbif_authomatic_matching}
biodiv_taxon_extended %>% 
  filter(
    acceptedname %in% c("Characium",
                        "Dicranochaete",
                        "Euzonus",
                        "Grandinia",
                        "Thorea") & 
      parentid_corrected %in% c(101415,
                                101420,
                                44193,
                                101473,
                                101408)) %>%
  select(id, 
         acceptedname, 
         scientificnameauthorship, 
         parentid, 
         parentid_corrected,
         everything())
```

These errors in GBIF match are very rare and occur because the match tries to refer to a non empty authorship, while the taxon has `scientificName` equal to `canonicalName`. Examples:  

```{r Thorea}
name_backbone("Thorea", rank = "genus", strict = TRUE, verbose = TRUE)$alternatives
```

```{r Euzonus}
name_backbone("Euzonus", rank = "genus", strict = TRUE, verbose = TRUE)$alternatives
```

```{r Grandinia}
name_backbone("Grandinia", rank = "genus", strict = TRUE, verbose = TRUE)$alternatives
```

The only way to solve these issues in the future is adding the full taxonomic tree as additional columns, so that we can better match to GBIF Backbone.

## Remove unused duplicates

We remove the unused duplicates discussed above:

```{r save_biodiv.taxon_without_unused_duplicates}
biodiv_taxon_without_unused_duplicates <- 
  biodiv_taxon %>%
  filter(!id %in% not_used_duplicate_names$id)
```

# Add `bruenvi_created` 

We add date of creation for new taxa:

```{r add_creation_date}
biodiv_taxon_without_unused_duplicates <- 
  biodiv_taxon %>%
  mutate(bruenvi_created = case_when(
    is.na(bruenvi_created) |  bruenvi_created == "" ~ as.character(Sys.Date()),
    TRUE ~ bruenvi_created
  ))
```

## Modify `bruenvi_modified`

Modified taxa:

1. taxa in `data/corrected_taxa.tsv`
2. taxa with modified parent ID


Read the corrected taxa:

```{r read_corrected_taxa}
corrected_taxa <-  read_tsv(
  here("references", "corrected_taxa.tsv"),
  na = "")
```

```{r modified_taxa}
taxonid_modified <-
  biodiv_taxon_extended %>%
  filter(parentid_corrected != parentid) %>%
  distinct(id) %>%
  bind_rows(corrected_taxa %>%
              select(id)) %>%
  distinct() %>%
  pull()
```

```{r modify_bruenvi_modified}
biodiv_taxon_without_unused_duplicates <- 
  biodiv_taxon_without_unused_duplicates %>%
  mutate(bruenvi_modified = case_when(
    id %in% taxonid_modified ~ as.character(Sys.Date()),
    TRUE ~ bruenvi_modified
  ))
```

Save taxa as UTF-8 (default) using pipe `|` as delimiter:

```{r save_biodiv_taxon_without_unused_duplicates}
write_delim(biodiv_taxon_without_unused_duplicates, 
            path = here("data", "processed", "biodiv.taxon.txt"), 
            na = "",
            delim = "|")
```
