---
title: "Match taxa with GBIF Taxonomy Backbone"
author:
  - Damiano Oldoni
  - Lien Reyserhove
  - Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

# Setup

## Load libraries

```{r load_libs}
library(odbc)       # To work with database
library(tidyverse)  # To do data science
library(here)       # To work with paths
library(inborutils) # To match taxa (rgbif called in background)
```

## Load access informations

Retrieve access informations from configuration file:

```{r get_access_infos}
ibge_bim <- config::get("ibge_bim")
```

# Read taxa

Read taxa from `./data/interim/taxa.tsv`:

```{r read_taxa_from_tsv_file}
taxa <- read_tsv(here("data", "interim", "taxa.tsv"),
                 col_types = cols(
                   .default = col_character(),
                   id = col_double(),
                   parentid = col_double(),
                   media = col_logical(),
                   speciesbeparentid = col_double(),
                   bruenvi_created = col_date(format = "%Y-%m-%d"),
                   bruenvi_created = col_date(format = "%Y-%m-%d"),
                   kingdom_id = col_double()
                   )
)
```

# Match taxa to GBIF Taxonomy Backbone

We match taxa to GBIF Taxonomy Backbone in three steps:

1. match by name with authorship + kingdom + rank
2. match by name without authorship + kingdom + rank
3. match by name without authorship + rank

## Match by `acceptedname` + `scientificnameauthorship`

We try to match by using `acceptedname` + `scientificnameauthorship` as name. We also provide kingdom and rank information to the match function. First, we prepare `taxa` for matching. In this step:

1. We create a new column, `name_for_gbif`, where we paste `acceptedname` and `scientificnameauthorship`, if present.
2. Recode the taxon ranks from `taxonranken` in a new column, `rank_for_gbif`, containing the correct GBIF terms.

```{r prepare_taxa_for_matching_backbone}
taxa <-
  taxa %>%
  mutate(name_for_gbif = if_else(
    !is.na(scientificnameauthorship),
    paste(acceptedname, scientificnameauthorship),
    acceptedname
  )) %>%
  mutate(rank_for_gbif = recode(taxonranken,
    "forma" = "form",
    "subforma" = "subform",
    "informal group" = NA_character_
  ))
```

We match the names in `name_for_gbif` with GBIF Backbone and retrieve the following GBIF fields:

```{r gbif_terms}
gbif_terms <- c("matchType",
                "confidence", 
                "rank", 
                "usageKey", 
                "scientificName",
                "canonicalName",
                "kingdom",
                "phylum",
                "order", 
                "class", 
                "family",
                "genus", 
                "species",
                "kingdomKey",
                "phylumKey",
                "classKey",
                "orderKey",
                "familyKey",
                "genusKey",
                "speciesKey",
                "status",
                "synonym", 
                "acceptedUsageKey")
```

We also use the columns containing kingdom information (`bim_kingdom`) and rank (`rank_for_gbif`) and we require a strict match, `strict = TRUE` to avoid fuzzy match with higher classification. **Note: This step takes long**:

```{r get_name_gbif}
taxa_gbif <-
  taxa %>%
  gbif_species_name_match(name = "name_for_gbif",
                          kingdom = "bim_kingdom", 
                          rank = "rank_for_gbif", 
                          gbif_terms = gbif_terms,
                          strict = TRUE)
```

We add the prefix `gbif_` to the columns coming from GBIF Backbone:

```{r add_gbif_prefix}
taxa_gbif <- 
  taxa_gbif %>%
  select(gbif_terms) %>%
  setNames(paste0('gbif_', names(.))) %>%
  bind_cols(taxa_gbif %>%
              select(-gbif_terms)) %>%
  select(-contains("gbif_"), everything())
```

Preview:

```{r overview_taxa_gbif}
taxa_gbif %>%
  head()
```

The column `gbif_matchType` informs us whether the match has been successful (`gbif_matchType` = `EXACT`), doubtful (`gbif_matchType` = `FUZZY`) or failed (`gbif_matchType` = `NONE`). The column  `confidence` is also informative. Number of taxa for each value of `gbif_matchType`:

```{r overview_n_match}
taxa_gbif %>%
  group_by(gbif_matchType) %>%
  count()
```

We create a new data.frame, `taxa_no_exact_match`, for taxa without exact match:

```{r taxa_no_exact_match}
taxa_no_exact_match <-
  taxa_gbif %>%
  filter(gbif_matchType != "EXACT") %>%
  arrange(id)
```

Preview (only relevant columns shown):

```{r not_matched_taxa}
taxa_no_exact_match %>%
  select(-c(speciesbeid, colid, vernacularnamefr, 
            vernacularnamenl, parentid, distributionstatuscode, 
            media, speciesbeparentid, bruenvi_created, bruenvi_author,
            bruenvi_modified, kingdom_id)) %>%
  head()
```

## Match without `scientificnameauthorship` only

Some of the taxa are not matched due to deviating author spellings, e.g.

```{r example_no_match_due_to_author_spelling_deviation}
taxa_no_exact_match %>%
  filter(name_for_gbif == "Neosartorya fischeri var. glabra (Fennell & Raper) Malloch & Cain")
```

while in GBIF Backbone:

```{r show_version_on_gbif}
rgbif::name_backbone(name = "`Neosartorya fischeri var. glabra Fennell & Raper, 1973`") %>% as_tibble()
```

Matching without authors can improve matching:

```{r match_acceptedname_only}
taxa_without_author <-
  taxa_no_exact_match %>%
  select(names(taxa)) %>%
  gbif_species_name_match(name = "acceptedname",
                          kingdom = "bim_kingdom", 
                          rank = "rank_for_gbif",
                          gbif_terms = gbif_terms,
                          strict = TRUE)
# add prefix gbif_
taxa_without_author <-
  taxa_without_author %>% 
  select(gbif_terms) %>%
  setNames(paste0('gbif_', names(.))) %>%
  bind_cols(taxa_without_author %>%
              select(-gbif_terms)) %>%
  select(-contains("gbif_"), everything())
```

Overview match:

```{r summarize_new_match}
taxa_without_author %>%
  group_by(gbif_matchType) %>%
  count()
```

As some of the matches might be wrong, we update `gbif_matchType` to `EXACT_WITHOUT_AUTHOR`, `FUZZY_WITHOUT_AUTHOR`:

```{r flag_matchType}
taxa_without_author <-
  taxa_without_author %>%
  mutate(gbif_matchType = case_when(
    gbif_matchType == "EXACT" ~ "EXACT_WITHOUT_AUTHOR",
    gbif_matchType == "FUZZY" ~ "FUZZY_WITHOUT_AUTHOR",
    TRUE ~ gbif_matchType)
)
```

Include the results of this second match attempt to `taxa_gbif`:

```{r join_taxa_without_author}
taxa_gbif <-
  taxa_gbif %>%
  anti_join(taxa_without_author,
            by = "id") %>%
  bind_rows(taxa_without_author) %>%
  arrange(id)
```

Overview:

```{r overview_matchType}
taxa_gbif %>%
  group_by(gbif_matchType) %>%
  count()
```

## Match without authorship and without kingdom

Due to errors in parent IDs, some taxa are not matched because they are linked to the wrong kingdom. Examples, the phylum `Xanthophyta`:

```{r example_Xanthophyta}
taxa_gbif %>% filter(acceptedname == "Xanthophyta")
```

Its `bim_kingdom` is based on parentid `45975`, which refers to the internal id of acceptedname `Plantae`, while the kingdom according to GBIF Backbone is  `Chromista`:

```{r example_Xanthophyta_GBIF_Backbone}
taxa_gbif %>% 
  filter(acceptedname == "Xanthophyta") %>%
  select(acceptedname) %>%
  gbif_species_name_match(name = "acceptedname") %>%
  select(acceptedname, kingdom) %>%
  rename(gbif_kingdom = kingdom)
```

The taxa with `gbif_matchType` not equal to `EXACT` or `EXACT_WITHOUT_AUTHOR` are matched again. As third and final matching attempt, we try matching taxa without `scientificnameauthorship` and without passing kingdom as parameter. Please note that rank is always given to the match function.

```{r match_without_kingdom_acceptedname_only}
taxa_without_author_and_kingdom <-
  taxa_gbif %>%
  filter(!gbif_matchType %in% c("EXACT_WITHOUT_AUTHOR", 
                           "EXACT")) %>%
  select(names(taxa)) %>%
  gbif_species_name_match(name = "acceptedname",
                          rank = "rank_for_gbif",
                          gbif_terms = gbif_terms,
                          strict = TRUE)
taxa_without_author_and_kingdom <-
  taxa_without_author_and_kingdom %>%
  select(gbif_terms) %>%
  setNames(paste0('gbif_', names(.))) %>%
  bind_cols(taxa_without_author_and_kingdom %>%
              select(-gbif_terms)) %>%
  select(-contains("gbif_"), everything())
```

Overview match:

```{r summarize_match_without_kingdom_acceptedname}
taxa_without_author_and_kingdom %>%
  group_by(gbif_matchType) %>%
  count()
```

As some of the matches might be wrong, we update `gbif_matchType` to `EXACT_WITHOUT_AUTHOR_AND_KINGDOM`, `FUZZY_WITHOUT_AUTHOR_AND_KINGDOM`:

```{r flag_matchType_without_author_and_kingdom}
taxa_without_author_and_kingdom <-
  taxa_without_author_and_kingdom %>%
  mutate(gbif_matchType = case_when(
    gbif_matchType == "EXACT" ~ "EXACT_WITHOUT_AUTHOR_AND_KINGDOM",
    gbif_matchType == "FUZZY" ~ "FUZZY_WITHOUT_AUTHOR_AND_KINGDOM",
    TRUE ~ gbif_matchType)
)
```

Include the results of this third match attempt to `taxa_gbif`:

```{r join_taxa_without_author_and_kingdom}
taxa_gbif <-
  taxa_gbif %>%
  anti_join(taxa_without_author_and_kingdom,
            by = "id") %>%
  bind_rows(taxa_without_author_and_kingdom) %>%
  arrange(id)
```

Overview  `gbif_matchType`:

```{r overview_matchType_3}
taxa_gbif %>%
  group_by(gbif_matchType) %>%
  count()
```

## Get number of occurrences per taxa

It is interesting to know how many occurrences are linked to the unmatched taxa. For example, taxa linked to many occurrences could be checked before other ones, thus prioritizing the work of the taxonomists. We retrieve this information by reading the occurrences from database.

### Connect to database

Connect to database:

```{r connect_to_db}
conn <- dbConnect(odbc(), 
                  driver = ibge_bim$driver,
                  server = ibge_bim$server,
                  database = ibge_bim$database,
                  port = ibge_bim$port,
                  uid = ibge_bim$uid,
                  pwd = ibge_bim$pwd,
                  encoding = "latin1"
)
```

### Get occurrences from table `occurence`

Table `occurence` contains occurrence data. It is linked to taxa in table `taxon` via intermediate table `identifiablespecies`, which takes care of synonym relations. We retrieve first the column `identifiablespeciesid` from table `occurence`:

```{r get_identifiable_species_ids}
identifiable_species_id <- 
  dbGetQuery(conn, "SELECT identifiablespeciesid FROM biodiv.occurence") %>%
  as_tibble()
```

Number of occurrences:

```{r}
nrow(identifiable_species_id)
```

Number of unique IDs, i.e. number of taxa (synonyms or accepted by BIM) linked to at least one occurrence:

```{r n_taxa_in_occs}
identifiable_species_id %>%
  distinct(identifiablespeciesid) %>%
  nrow()
```

The IDs in `identifiable_species_id` are linked to column `id` in table `identifiablespecies`:
 
```{r get_identifiable_species}
identifiable_species <- 
  dbGetQuery(conn, "SELECT * FROM biodiv.identifiablespecies") %>%
  as_tibble()
```

Preview:

```{r overview_identifiable_species}
head(identifiable_species)
```

Column `taxonid` contains the taxon identifiers corresponding to column `id` of table `biodiv.taxon`.

Number of taxa in this intermediate table:

```{r n_taxa_identifiable_species}
nrow(identifiable_species)
```

Number of distinct accepted taxa:

```{r n_unique_taxa}
identifiable_species %>%
  distinct(taxonid) %>%
  nrow()
```

Taxa linked to multiple IDs, thus collecting synonyms:

```{r multiple_IDs}
identifiable_species %>%
  group_by(taxonid) %>%
  count() %>%
  filter(n > 1) %>%
  left_join(identifiable_species, by = "taxonid") %>%
  select(names(identifiable_species))
```

Notice how these rows can contain synonyms only. 

Example: taxa linked to  `taxonid` 79:

```{r example_synonyms_only}
identifiable_species %>%
  filter(taxonid == 79) %>%
  select(id, taxonid, taxonomicstatusen)
```

The acepted taxon is the following:

```{r accepted_taxon_exmaple} 
taxa_gbif %>%
  filter(id == 79) %>%
  select(id, acceptedname, scientificnameauthorship, taxonomicstatusen)
```

There also accepted taxa not present in `identifiable_species`:

```{r taxa_not_in_identifiable_species}
taxa_gbif %>%
  filter(!id %in% identifiable_species$taxonid)
```

We add `taxonid` column to `identifiable_species_id` data.frame:

```{r add_taxonid}
identifiable_species_id <-
  identifiable_species_id %>%
  left_join(identifiable_species %>% select(id, taxonid),
            by = c("identifiablespeciesid" = "id"))
```

So, we got a data.frame containing the accepted taxon ID of each occurrence.

Number of acceted taxa linked to at least one occurrence:

```{r n_accepted_taxa}
identifiable_species_id %>%
  distinct(taxonid) %>%
  nrow()
```

### Close connection

Close connection:

```{r close_connection_occs}
dbDisconnect(conn)
```

### Add number of occurrences to taxonomic data

We add a new column called `n_occs` containing the number of occurrences linked to each taxon:

```{r add_n_occs}
taxa_gbif <- 
  identifiable_species_id %>%
  group_by(taxonid) %>%
  count() %>%
  ungroup() %>%
  rename(n_occs = n) %>%
  right_join(taxa_gbif,
             by = c("taxonid" = "id")) %>%
  rename(id = taxonid) %>%
  arrange(id)
```

Number of taxa with/without occurrences:

```{r n_taxa_with_without_occs}
taxa_gbif %>%
  group_by(!is.na(n_occs)) %>%
  count() %>%
  rename("has_occs" = "!is.na(n_occs)")
```

### Number of occurrences of unmatched taxa

In particular, for unmatched taxa (`gbif_matchType` not equal to `EXACT`):

```{r n_occs_unmatched_taxa}
taxa_gbif_unmatched <- 
  taxa_gbif %>%
  filter(gbif_matchType != "EXACT") %>%
  select(acceptedname,
         scientificnameauthorship,
         n_occs, 
         gbif_matchType, 
         gbif_confidence, 
         gbif_status, 
         gbif_usageKey,
         gbif_scientificName,
         gbif_canonicalName,
         gbif_rank,
         gbif_kingdom,
         gbif_phylum,
         gbif_class,
         gbif_order,
         gbif_family,
         gbif_genus,
         gbif_species,
         everything())
# preview
taxa_gbif_unmatched %>%
  head()
```

Number of unmatched taxa linked to at least one occurrence _or_ none at all **and** with _or_ without parent ID:

```{r unmatched_taxa_count_n_occs}
taxa_gbif_unmatched %>%
  group_by(!is.na(parentid), !is.na(n_occs)) %>%
  count() %>%
  rename("has_parentid" = "!is.na(parentid)",
         "has_occs" = "!is.na(n_occs)"
)
```

Experts should focus first on taxa without parent ID, `has_parentid = FALSE`. Among these ones priority should be given to taxa with occurrences, `has_occs = TRUE`. To do so, we arrange the taxa by absence of parent ID and decreasing number of occurrences:

```{r taxa_gbif_unmatched_arrange_parentid_n_occs}
taxa_gbif_unmatched <- 
  taxa_gbif_unmatched %>%
  arrange(!is.na(parentid), n_occs)
```

Finally, it is also important to show the rank of the unmatched taxa:

```{r rank_of_unmatched_taxa}
taxa_gbif_unmatched %>%
  group_by(rank_for_gbif) %>%
  count()
```

# Save taxa

We save the taxa with the added information from GBIF Taxonomy Backbone:

```{r save_taxa_gbif}
write_tsv(taxa_gbif,
          path = here::here("data", "interim", "taxa_gbif.tsv"),
          na = "")
```

We save the subset of the unmatched taxa in a separate file for verification.

```{r save_unmatched_taxa}
write_tsv(taxa_gbif_unmatched,
          path = here::here("data", "interim", "unmatched_taxa_gbif.tsv"),
          na = "")
```
