---
title: "Match taxa to GBIF Taxonomy Backbone"
author:
  - Damiano Oldoni
  - Lien Reyserhove
  - Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

# Setup

## Load libraries

```{r load_libs}
library(odbc)       # To work with database
library(tidyverse)  # To do data science
library(here)       # To work with paths
library(inborutils) # To match taxa (rgbif called in background)
```

## Load access informations

Retrieve access informations from configuration file:

```{r get_access_infos}
ibge_bim <- config::get("ibge_bim")
```

# Read taxa

Read taxa from `./data/interim/taxa.tsv`:

```{r read_taxa_from_tsv_file}
taxa <- read_tsv(here("data", "interim", "taxa.tsv"),
                 col_types = cols(
                   .default = col_character(),
                   id = col_double(),
                   parentid = col_double(),
                   media = col_logical(),
                   speciesbeparentid = col_double(),
                   bruenvi_created = col_date(format = "%Y-%m-%d"),
                   bruenvi_created = col_date(format = "%Y-%m-%d"),
                   kingdom_id = col_double()
                   )
)
```

# Match taxa to GBIF Taxonomy Backbone

We match taxa to GBIF Taxonomy Backbone in three steps:

1. match by scientific name with authorship + kingdom + rank
2. match by scientific name + kingdom + rank
3. match by scientific name + rank

## Match by `acceptedname` + `scientificnameauthorship` + `kingdom` + `rank`

We try to match by using `acceptedname` + `scientificnameauthorship` as name. We also provide kingdom and rank information to the match function. First, we prepare `taxa` for matching. In this step:

1. We create a new column, `name_for_gbif`, where we paste `acceptedname` and `scientificnameauthorship`, if present.
2. Recode the taxon ranks from `taxonranken` in a new column, `rank_for_gbif`, containing the correct GBIF terms.

```{r prepare_taxa_for_matching_backbone}
taxa <-
  taxa %>%
  mutate(name_for_gbif = if_else(
    !is.na(scientificnameauthorship),
    paste(acceptedname, scientificnameauthorship),
    acceptedname
  )) %>%
  mutate(rank_for_gbif = recode(taxonranken,
    "forma" = "form",
    "subforma" = "subform",
    "informal group" = NA_character_
  ))
```

We match the names in `name_for_gbif` with GBIF Backbone and retrieve the following GBIF fields:

```{r gbif_terms}
gbif_terms <- c("matchType",
                "confidence", 
                "rank", 
                "usageKey", 
                "scientificName",
                "canonicalName",
                "kingdom",
                "phylum",
                "order", 
                "class", 
                "family",
                "genus", 
                "species",
                "kingdomKey",
                "phylumKey",
                "classKey",
                "orderKey",
                "familyKey",
                "genusKey",
                "speciesKey",
                "status",
                "synonym", 
                "acceptedUsageKey")
```

We also use the columns containing kingdom information (`bim_kingdom`) and rank (`rank_for_gbif`) and we require a strict match, `strict = TRUE` to avoid fuzzy match with higher classification. **Note: This step takes long**:

```{r get_name_gbif, cache=TRUE}
taxa_gbif <-
  taxa %>%
  gbif_species_name_match(name = "name_for_gbif",
                          kingdom = "bim_kingdom", 
                          rank = "rank_for_gbif", 
                          gbif_terms = gbif_terms,
                          strict = TRUE)
```

We add the prefix `gbif_` to the columns coming from GBIF Backbone:

```{r add_gbif_prefix}
taxa_gbif <- 
  taxa_gbif %>%
  select(gbif_terms) %>%
  setNames(paste0('gbif_', names(.))) %>%
  bind_cols(taxa_gbif %>%
              select(-gbif_terms)) %>%
  select(-contains("gbif_"), everything())
```

Preview:

```{r overview_taxa_gbif}
taxa_gbif %>%
  head()
```

The column `gbif_matchType` informs us whether the match has been successful (`gbif_matchType` = `EXACT`), doubtful (`gbif_matchType` = `FUZZY`) or failed (`gbif_matchType` = `NONE`). The column  `confidence` is also informative. Number of taxa for each value of `gbif_matchType`:

```{r overview_n_match}
taxa_gbif %>%
  group_by(gbif_matchType) %>%
  count()
```

We create a new data.frame, `taxa_no_exact_match`, for taxa without exact match:

```{r taxa_no_exact_match}
taxa_no_exact_match <-
  taxa_gbif %>%
  filter(gbif_matchType != "EXACT") %>%
  arrange(id)
```

Preview (only relevant columns shown):

```{r not_matched_taxa}
taxa_no_exact_match %>%
  select(-c(speciesbeid, colid, vernacularnamefr, 
            vernacularnamenl, parentid, distributionstatuscode, 
            media, speciesbeparentid, bruenvi_created, bruenvi_author,
            bruenvi_modified, kingdom_id)) %>%
  head()
```

## Match by `acceptedname` + `kingdom` + `rank`

Some of the taxa are not matched due to deviating author spellings, e.g.

```{r example_no_match_due_to_author_spelling_deviation}
taxa_no_exact_match %>%
  filter(name_for_gbif == "Neosartorya fischeri var. glabra (Fennell & Raper) Malloch & Cain")
```

while in GBIF Backbone:

```{r show_version_on_gbif}
rgbif::name_backbone(name = "`Neosartorya fischeri var. glabra Fennell & Raper, 1973`") %>% as_tibble()
```

Matching without authors can improve matching:

```{r match_acceptedname_only, cache=TRUE}
taxa_without_author <-
  taxa_no_exact_match %>%
  select(names(taxa)) %>%
  gbif_species_name_match(name = "acceptedname",
                          kingdom = "bim_kingdom", 
                          rank = "rank_for_gbif",
                          gbif_terms = gbif_terms,
                          strict = TRUE)
# add prefix gbif_
taxa_without_author <-
  taxa_without_author %>% 
  select(gbif_terms) %>%
  setNames(paste0('gbif_', names(.))) %>%
  bind_cols(taxa_without_author %>%
              select(-gbif_terms)) %>%
  select(-contains("gbif_"), everything())
```

Overview match:

```{r summarize_new_match}
taxa_without_author %>%
  group_by(gbif_matchType) %>%
  count()
```

As some of the matches might be wrong, we update `gbif_matchType` to `EXACT_WITHOUT_AUTHOR`, `FUZZY_WITHOUT_AUTHOR`:

```{r flag_matchType}
taxa_without_author <-
  taxa_without_author %>%
  mutate(gbif_matchType = case_when(
    gbif_matchType == "EXACT" ~ "EXACT_WITHOUT_AUTHOR",
    gbif_matchType == "FUZZY" ~ "FUZZY_WITHOUT_AUTHOR",
    TRUE ~ gbif_matchType)
)
```

Include the results of this second match attempt to `taxa_gbif`:

```{r join_taxa_without_author}
taxa_gbif <-
  taxa_gbif %>%
  anti_join(taxa_without_author,
            by = "id") %>%
  bind_rows(taxa_without_author) %>%
  arrange(id)
```

Overview:

```{r overview_matchType}
taxa_gbif %>%
  group_by(gbif_matchType) %>%
  count()
```

## Match by `acceptedname` + `rank`

Due to errors in parent IDs, some taxa are not matched because they are linked to the wrong kingdom. Examples, the phylum `Xanthophyta`:

```{r example_Xanthophyta}
taxa_gbif %>% filter(acceptedname == "Xanthophyta")
```

Its `bim_kingdom` is based on parentid `45975`, which refers to the internal id of acceptedname `Plantae`, while the kingdom according to GBIF Backbone is  `Chromista`:

```{r example_Xanthophyta_GBIF_Backbone}
taxa_gbif %>% 
  filter(acceptedname == "Xanthophyta") %>%
  select(acceptedname) %>%
  gbif_species_name_match(name = "acceptedname") %>%
  select(acceptedname, kingdom) %>%
  rename(gbif_kingdom = kingdom)
```

The taxa with `gbif_matchType` not equal to `EXACT` or `EXACT_WITHOUT_AUTHOR` are matched again. As third and final matching attempt, we try matching taxa without `scientificnameauthorship` and without passing kingdom as parameter. Please note that rank is always given to the match function.

```{r match_without_kingdom_acceptedname_only, cache=TRUE}
taxa_without_author_and_kingdom <-
  taxa_gbif %>%
  filter(!gbif_matchType %in% c("EXACT_WITHOUT_AUTHOR", 
                           "EXACT")) %>%
  select(names(taxa)) %>%
  gbif_species_name_match(name = "acceptedname",
                          rank = "rank_for_gbif",
                          gbif_terms = gbif_terms,
                          strict = TRUE)
taxa_without_author_and_kingdom <-
  taxa_without_author_and_kingdom %>%
  select(gbif_terms) %>%
  setNames(paste0('gbif_', names(.))) %>%
  bind_cols(taxa_without_author_and_kingdom %>%
              select(-gbif_terms)) %>%
  select(-contains("gbif_"), everything())
```

Overview match:

```{r summarize_match_without_kingdom_acceptedname}
taxa_without_author_and_kingdom %>%
  group_by(gbif_matchType) %>%
  count()
```

As some of the matches might be wrong, we update `gbif_matchType` to `EXACT_WITHOUT_AUTHOR_AND_KINGDOM`, `FUZZY_WITHOUT_AUTHOR_AND_KINGDOM`:

```{r flag_matchType_without_author_and_kingdom}
taxa_without_author_and_kingdom <-
  taxa_without_author_and_kingdom %>%
  mutate(gbif_matchType = case_when(
    gbif_matchType == "EXACT" ~ "EXACT_WITHOUT_AUTHOR_AND_KINGDOM",
    gbif_matchType == "FUZZY" ~ "FUZZY_WITHOUT_AUTHOR_AND_KINGDOM",
    TRUE ~ gbif_matchType)
)
```

Include the results of this third match attempt to `taxa_gbif`:

```{r join_taxa_without_author_and_kingdom}
taxa_gbif <-
  taxa_gbif %>%
  anti_join(taxa_without_author_and_kingdom,
            by = "id") %>%
  bind_rows(taxa_without_author_and_kingdom) %>%
  arrange(id)
```

Overview  `gbif_matchType`:

```{r overview_matchType_3}
taxa_gbif %>%
  group_by(gbif_matchType) %>%
  count()
```

## Get number of occurrences per taxa

It is interesting to know how many occurrences are linked to the unmatched taxa: taxa linked to many occurrences should be checked before other ones, thus prioritizing the work of the taxonomists. We retrieve this information by reading the occurrences from text file:

```{r read_occurrences}
occurrences <- read_tsv(here("data", "input", "occurrences.tsv"), 
                        na = "",
                        col_types = cols(
                          .default = col_character(),
                          id = col_double(),
                          locationid = col_double(),
                          identificationid = col_double(),
                          identifiablespeciesid = col_double(),
                          observerid = col_double(),
                          eventid = col_double(),
                          individualcount = col_double(),
                          individualcountestimationid = col_double(),
                          dateentered = col_date(format = ""),
                          bbd_id = col_double(),
                          bruenvi_created = col_date(format = "%Y-%m-%d"),
                          bruenvi_modified = col_date(format = "%Y-%m-%d"))
)
```

The link to `taxa` has been made by intermediate table `identifiablespecies` containing all identifiable species and linking synonyms to accepted taxa. We read it by file:

```{r read_identifiable_species}
identifiable_species <- read_tsv(here("data", 
                                      "input", 
                                      "identifiable_species.tsv"), 
                                 na = "",
                                 col_types = cols(
                                   .default = col_character(),
                                   id = col_double(),
                                   taxonid = col_double(),
                                   bruenvi_created = col_date(format = "%Y-%m-%d"),
                                   bruenvi_modified = col_date(format = "%Y-%m-%d"))
)
```

Link observations to taxonomic ID of accepted taxa in `taxa`:

```{r link_to_taxa}
occurrences <- 
  occurrences %>%
  left_join(identifiable_species %>%
              select(id, taxonid),
            by = c("identifiablespeciesid" = "id"))
```

### Add number of occurrences to taxonomic data

We add a new column called `n_occs` containing the number of occurrences linked to each taxon:

```{r add_n_occs}
taxa_gbif <- 
  occurrences %>%
  group_by(taxonid) %>%
  count() %>%
  ungroup() %>%
  rename(n_occs = n) %>%
  right_join(taxa_gbif,
             by = c("taxonid" = "id")) %>%
  rename(id = taxonid) %>%
  arrange(id)
```

Number of taxa with/without occurrences:

```{r n_taxa_with_without_occs}
taxa_gbif %>%
  group_by(!is.na(n_occs)) %>%
  count() %>%
  rename("has_occs" = "!is.na(n_occs)")
```

### Number of occurrences of unmatched taxa

In particular, for unmatched taxa (`gbif_matchType` not equal to `EXACT`):

```{r n_occs_unmatched_taxa}
taxa_gbif_unmatched <- 
  taxa_gbif %>%
  filter(gbif_matchType != "EXACT") %>%
  select(acceptedname,
         scientificnameauthorship,
         n_occs, 
         gbif_matchType, 
         gbif_confidence, 
         gbif_status, 
         gbif_usageKey,
         gbif_scientificName,
         gbif_canonicalName,
         gbif_rank,
         gbif_kingdom,
         gbif_phylum,
         gbif_class,
         gbif_order,
         gbif_family,
         gbif_genus,
         gbif_species,
         everything())
# preview
taxa_gbif_unmatched %>%
  head()
```

Number of unmatched taxa linked to at least one occurrence _or_ none at all **and** with _or_ without parent ID:

```{r unmatched_taxa_count_n_occs}
taxa_gbif_unmatched %>%
  group_by(!is.na(parentid), !is.na(n_occs)) %>%
  count() %>%
  rename("has_parentid" = "!is.na(parentid)",
         "has_occs" = "!is.na(n_occs)"
)
```

Experts should focus first on taxa without parent ID, `has_parentid = FALSE`. Among these ones priority should be given to taxa with occurrences, `has_occs = TRUE`. To do so, we arrange the taxa by absence of parent ID and decreasing number of occurrences:

```{r taxa_gbif_unmatched_arrange_parentid_n_occs}
taxa_gbif_unmatched <- 
  taxa_gbif_unmatched %>%
  arrange(!is.na(parentid), n_occs)
```

Finally, it is also important to show the rank of the unmatched taxa:

```{r rank_of_unmatched_taxa}
taxa_gbif_unmatched %>%
  group_by(rank_for_gbif) %>%
  count()
```

# Save taxa

We save the taxa with the added information from GBIF Taxonomy Backbone:

```{r save_taxa_gbif}
write_tsv(taxa_gbif,
          path = here::here("data", "interim", "taxa_gbif.tsv"),
          na = "")
```

We save the subset of the unmatched taxa in a separate file for verification.

```{r save_unmatched_taxa}
write_tsv(taxa_gbif_unmatched,
          path = here::here("data", "interim", "unmatched_taxa_gbif.tsv"),
          na = "")
```
