---
title: "Match taxa to GBIF Backbone after review"
author:
  - Damiano Oldoni
  - Lien Reyserhove
  - Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

# Setup

## Load libraries

```{r match_corrected_taxa_to_gbif-load_libs}
library(tidyverse)  # To do data science
library(tidylog)    # To provide feedback on dplyr functions
library(here)       # To work with paths
library(inborutils) # To match taxa (rgbif called in background)
```

# Read taxonomic data

## Read taxa after review

We read the taxa with the corrected name, rank and kingdom information:

```{r match_corrected_taxa_to_gbif-read_flagged_taxa}
taxa_after_correction <- read_tsv(
  here("data", "interim", "corrected_flagged_taxa.tsv"),
  col_types = cols(
    .default = col_character(),
    id = col_double(),
    parentid = col_double(),
    media = col_logical(),
    speciesbeparentid = col_double(),
    bruenvi_created = col_date(format = "%Y-%m-%d"),
    bruenvi_modified = col_date(format = "%Y-%m-%d"),
    corrected = col_logical())
)
```

The column `corrected` acts as a flag to indicate whether experts corrected any of  `acceptedname`, `scientificnameauthorship`, `bim_kingdom` or `taxonranken`. Number of taxa after correction:

```{r match_corrected_taxa_to_gbif-n_corrected_in_use}
taxa_after_correction %>%
  group_by(corrected) %>%
  summarize(n_taxa = n())
```

Notice that the great majority of taxa don't need to be corrected!

## Read taxa with GBIF information

We read the full `taxa` after applying the first attempts to match taxa  with GBIF Taxonomy Backbone:

```{r match_corrected_taxa_to_gbif-read_taxa_table_with_GBIF}
taxa_gbif <- read_tsv(here("data", "interim", "taxa_gbif.tsv"),
                      col_types = cols(
                        .default = col_character(),
                        id = col_double(),
                        n_occs = col_double(),
                        parentid = col_double(),
                        media = col_logical(),
                        speciesbeparentid = col_double(),
                        bruenvi_created = col_date(format = "%Y-%m-%d"),
                        bruenvi_modified = col_date(format = "%Y-%m-%d"),
                        kingdom_id = col_double(),
                        gbif_usageKey = col_double(),
                        gbif_confidence = col_double(),
                        gbif_kingdomKey = col_double(),
                        gbif_phylumKey = col_double(),
                        gbif_classKey = col_double(),
                        gbif_orderKey = col_double(),
                        gbif_familyKey = col_double(),
                        gbif_genusKey = col_double(),
                        gbif_synonym = col_logical(),
                        gbif_acceptedUsageKey = col_double(),
                        gbif_speciesKey = col_double())
)
```

The columns with prefix `gbif_` come from GBIF. The column `gbif_matchType` is particularly important as it indicates the type of match:

```{r match_corrected_taxa_to_gbif-type_of_match}
taxa_gbif %>%
  distinct(gbif_matchType)
```

# Add GBIF information to the corrected taxa in use 

First, we add the columns coming from GBIF to `taxa_after_correction`:

```{r match_corrected_taxa_to_gbif-add_GBIF_info_to_taxa_after_correction}
taxa_after_correction <-
  taxa_after_correction %>%
  left_join(taxa_gbif %>%
              select(id, starts_with("gbif_")),
            by = "id")
```

Type of match of taxa with `corrected` equal to `TRUE`:

```{r match_corrected_taxa_to_gbif-n_taxa_corrected_is_true}
corrected_taxa <-
  taxa_after_correction %>%
  filter(corrected == TRUE)
corrected_taxa %>%
  group_by(gbif_matchType) %>%
  count()
```

After being corrected, these taxa can be matched again to GBIF Backbone. 

## Prepare taxa for match to GBIF Backbone

### Name

We paste `acceptedname` and `scientificnameauthorship` together in column `name_for_gbif`, which will be used for match:

```{r match_corrected_taxa_to_gbif-prepare_name}
corrected_taxa <-
  corrected_taxa %>%
  select(-starts_with("gbif")) %>%
  select(-ends_with("gbif")) %>%
  mutate(name_for_gbif = if_else(
    !is.na(scientificnameauthorship),
    paste(acceptedname, scientificnameauthorship),
    acceptedname
  ))
```

### Rank

Map rank to GBIF vocabulary:

```{r match_corrected_taxa_to_gbif-corrected_taxa_rank_for_gbif}
corrected_taxa <-
  corrected_taxa %>%
  mutate(rank_for_gbif = recode(taxonranken,
    "forma" = "form",
    "subforma" = "subform",
    "informal group" = NA_character_,
    "division" = NA_character_
  ))
```

## Match to GBIF Backbone

### GBIF terms

GBIF terms we are interested to:

```{r match_corrected_taxa_to_gbif-gbif_terms}
gbif_terms <- c("matchType",
                "confidence", 
                "rank", 
                "usageKey", 
                "scientificName",
                "canonicalName",
                "kingdom",
                "phylum",
                "order", 
                "class", 
                "family",
                "genus", 
                "species",
                "kingdomKey",
                "phylumKey",
                "classKey",
                "orderKey",
                "familyKey",
                "genusKey",
                "speciesKey",
                "status",
                "synonym", 
                "acceptedUsageKey")
```

### Match

We match by name using also columns containing kingdom information (`bim_kingdom`) and rank (`rank_for_gbif`). We require a strict match, `strict = TRUE` to avoid fuzzy match with higher classification.

```{r match_corrected_taxa_to_gbif-get_name_gbif, cache=TRUE}
corrected_taxa_gbif <-
  corrected_taxa %>%
  gbif_species_name_match(name = "name_for_gbif",
                          kingdom = "bim_kingdom", 
                          rank = "rank_for_gbif", 
                          gbif_terms = gbif_terms,
                          strict = TRUE)
```

We add the prefix `gbif_` to the columns coming from GBIF Backbone:

```{r match_corrected_taxa_to_gbif-add_gbif_prefix}
cols_from_gbif <-
colnames(corrected_taxa_gbif)[
  colnames(corrected_taxa_gbif) %in% gbif_terms]
colnames(corrected_taxa_gbif)[
  colnames(corrected_taxa_gbif) %in% gbif_terms] <-
  paste0("gbif_", cols_from_gbif)
```

Overview match:

```{r match_corrected_taxa_to_gbif-overview_match}
corrected_taxa_gbif %>%
  group_by(gbif_matchType) %>%
  count()
```

Unmatched taxa after review:

```{r match_corrected_taxa_to_gbif-taxa_after_correction}
corrected_taxa_gbif %>%
  filter(gbif_matchType != "EXACT")
```

Notice how the rank of most of the unmached taxa is one of: subclass, division, subphylum, subclass, subfamily, superfamily. Such ranks are not valid in GBIF Backbone (see issue [#33](https://github.com/inbo/ibge-bim-species/issues/33)).

## Add retrieved GBIF information to taxa

Now that the corrected taxa have been matched to GBIF Backbone and that we have retrieved related taxonomic information from GBIF, we can merge them to the other (used) taxa which didn't need any correction:

```{r match_corrected_taxa_to_gbif-merge_corrected_taxa_to_taxa_gbif}
taxa_final_match_gbif <-
  taxa_after_correction %>%
  filter(corrected == FALSE) %>%
  bind_rows(corrected_taxa_gbif) %>%
  arrange(id)
```

# Save taxa with GBIF Backbone information

We save the taxa with added taxonomic information from GBIF Backbone:

```{r match_corrected_taxa_to_gbif-save_taxa_after_final_match_gbif}
write_tsv(taxa_final_match_gbif, 
          path = here("data", "interim", "taxa_final_match_gbif.tsv"),
          na = "")
```
