---
title: "Apply corrections on names, authors and rank"
author:
  - Damiano Oldoni
  - Lien Reyserhove
  - Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

# Setup

## Load libraries

```{r apply_corrections_names_by_ref_file-load_libs}
library(tidyverse)  # To do data science
library(tidylog)    # To provide feedback on dplyr functions
library(here)       # To work with paths
```

# Read taxonomic data

## Read taxa

We read taxa which we will further use to build the improved `taxa` table. In [previous pipeline](https://github.com/inbo/ibge-bim-species/blob/master/src/parentid_2_detect_used_taxa.Rmd) we added flag `used` to indicate whether the taxon is used in higher taxonomy or in linked tables.

```{r apply_corrections_names_by_ref_file-read_flagged_taxa}
flagged_taxa <- read_tsv(here("data", "interim", "flagged_taxa.tsv"),
                      col_types = cols(
                        .default = col_character(),
                        id = col_double(),
                        parentid = col_double(),
                        media = col_logical(),
                        speciesbeparentid = col_double(),
                        bruenvi_created = col_date(format = "%Y-%m-%d"),
                        bruenvi_modified = col_date(format = "%Y-%m-%d")))
```

## Read taxa with GBIF information

We read the full `taxa` after applying the first attempts to match taxa with GBIF Taxonomy Backbone:

```{r apply_corrections_names_by_ref_file-read_taxa_table_with_GBIF}
taxa_gbif <- read_tsv(here("data", "interim", "taxa_gbif.tsv"),
                      col_types = cols(
                        .default = col_character(),
                        id = col_double(),
                        n_occs = col_double(),
                        parentid = col_double(),
                        media = col_logical(),
                        speciesbeparentid = col_double(),
                        bruenvi_created = col_date(format = "%Y-%m-%d"),
                        bruenvi_modified = col_date(format = "%Y-%m-%d"),
                        kingdom_id = col_double(),
                        gbif_usageKey = col_double(),
                        gbif_confidence = col_double(),
                        gbif_kingdomKey = col_double(),
                        gbif_phylumKey = col_double(),
                        gbif_classKey = col_double(),
                        gbif_orderKey = col_double(),
                        gbif_familyKey = col_double(),
                        gbif_genusKey = col_double(),
                        gbif_synonym = col_logical(),
                        gbif_acceptedUsageKey = col_double(),
                        gbif_speciesKey = col_double())
)
```

## Read reference file

The unmatched taxa have been previously saved in a reference file called `corrected_taxa.tsv`. This file has been manually screened by an expert which added the corrected informations where needed:

```{r apply_corrections_names_by_ref_file-read_corrected_taxa.tsv}
corrected_taxa <- read_tsv(here("references", "corrected_taxa.tsv"),
                           na = "",
                           col_types = cols(
                             .default = col_character(),
                             id = col_double())
)
```

Preview:

```{r apply_corrections_names_by_ref_file-corrected_taxa_preview}
corrected_taxa %>%
  head()
```

The columns ending with suffix `_corrected` contain the corrected information where corrections have been applied. For example, in excerpt shown above we can see that the taxon zith canonical name `Teucrium chamaedrys germanicum` has been corrected as `Teucrium chamaedrys subsp. germanicum`, while `Russula heterophylla form. pseudoochroleuca` has been corrected as `Russula heterophylla f. pseudo-ochroleuca` and author `Romagn.` has been changed to `Romagn., 1962`.

In some cases we have to remove authorship. This was typically due to anomalous taxa within GBIF Backbone: issues to GBIF Portal have been created, see reference in column `remarks`. We used the `flag `EMPTY` to identify these names.

```{r apply_corrections_names_by_ref_file-show_flag_EMPTY}
corrected_taxa %>%
  filter(scientificnameauthorship_corrected == "EMPTY")
```

We set empty string where flag `EMPTY`:

```{r apply_corrections_names_by_ref_file-set_zero_length_string_where_flag_EMPTY}
corrected_taxa <-
  corrected_taxa %>%
  mutate(scientificnameauthorship_corrected = if_else(
    scientificnameauthorship_corrected == "EMPTY",
    "",
    scientificnameauthorship_corrected
  ))
```

# Apply corrections

Based on `corrected_taxa` we apply taxonomic corrections to `flagged_taxa`. Number of taxa whose taxonomic information will be corrected:

```{r apply_corrections_names_by_ref_file-n_flagged_taxa_to_correct}
corrected_taxa %>%
  inner_join(flagged_taxa,
            by = c("id", 
                   "acceptedname", 
                   "scientificnameauthorship", 
                   "taxonranken")) %>%
  nrow()
```

Apply corrections on columns `acceptedname`, `scientificnameauthorship`, `taxonranken` and `bim_kingdom`:

```{r apply_corrections_names_by_ref_file-correct_flagged_taxa}
flagged_taxa_after_correction <-
  flagged_taxa %>%
  
  # add kingdom based on BIM parent IDs
  left_join(taxa_gbif,
            by = names(flagged_taxa)[names(flagged_taxa) != "used"]) %>%
  select(-starts_with("gbif_")) %>%
  
  # add column with corrections
  left_join(corrected_taxa,
            by = c("id",
                   "acceptedname",
                   "scientificnameauthorship",
                   "taxonranken",
                   "bim_kingdom")) %>%
  
  # apply corrections
  mutate(acceptedname = if_else(!is.na(acceptedname_corrected),
                                acceptedname_corrected,
                                acceptedname),
         scientificnameauthorship = if_else(
           !is.na(scientificnameauthorship_corrected),
           scientificnameauthorship_corrected,
           scientificnameauthorship),
         taxonranken = if_else(!is.na(taxonranken_corrected),
                               taxonranken_corrected,
                               taxonranken),
         bim_kingdom = if_else(!is.na(bim_kingdom_corrected),
                               bim_kingdom_corrected,
                               bim_kingdom)) %>%
    
  # add column corrected as flag
  mutate(corrected = if_else(!is.na(acceptedname_corrected) |
                               !is.na(scientificnameauthorship_corrected) |
                               !is.na(taxonranken_corrected) | 
                               !is.na(bim_kingdom_corrected), TRUE, FALSE)) %>%
  
  # remove `*_corrected` columns
  select(-ends_with("_corrected"))
```

# Save corrected taxa

Save the taxa after applying corrections:

```{r apply_corrections_names_by_ref_file-save_flagged_taxa_after_correction}
write_tsv(flagged_taxa_after_correction, 
          path = here("data", "interim", "corrected_flagged_taxa.tsv"),
          na = "")
```
