---
title: "Review annexes"
author:
  - Lien Reyserhove
  - Damiano Oldoni
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Setup

## Load libraries

```{r load_libs}
library(tidyverse)      # To do data science
library(here)           # To work with paths
```

# Read data

We define first the annex codes we will use:

```{r annex_codes}
annexcode <- c("bxl_2.1",
               "bxl_2.2",
               "bxl_2.3",
               "bxl_2.4",
               "bxl_2.5",
               "bxl_3",
               "bxl_4",
               "bern_1",
               "bern_2",
               "bern_3",
               "bonn_1",
               "bonn_2",
               "hab_2",
               "hab_4")
```

This vector will be used later to identify the BIM annexes and the official annexes.

## Read BIM annex data

Import the BIM annex data with full taxonomic information:

```{r read_bim_annex_taxa}
bim_annex_taxa <- read_csv(
  here("data", "interim", "bim_annex_taxa_full_taxonomic_info.csv"),
  na = "",
  col_types = cols(
    .default = col_character(),
    taxonid = col_double(),
    parentid = col_double(),
    media = col_logical(),
    speciesbeparentid = col_double(),
    bruenvi_created = col_date(format = ""),
    bruenvi_modified = col_date(format = ""),
    used = col_logical(),
    n_occs = col_double(),
    kingdom_id = col_double(),
    corrected = col_logical(),
    gbif_confidence = col_double(),
    gbif_usageKey = col_double(),
    gbif_kingdomKey = col_double(),
    gbif_phylumKey = col_double(),
    gbif_classKey = col_double(),
    gbif_orderKey = col_double(),
    gbif_familyKey = col_double(),
    gbif_genusKey = col_double(),
    gbif_speciesKey = col_double(),
    gbif_synonym = col_logical(),
    gbif_acceptedUsageKey = col_double(),
    gbif_parentKey = col_double(),
    added_from_gbif = col_logical(),
    parentid_corrected = col_double())
)
```

Group BIM annexes by annex code:

```{r define_bim_annexcode}
bim_annexcode <- c(
  "BXL-ORD-2012_Annex II.1",
  "BXL-ORD-2012_Annex II.2",
  "BXL-ORD-2012_Annex II.3",
  "BXL-ORD-2012_Annex II.4",
  "BXL-ORD-2012_Annex II.5",
  "BXL-ORD-2012_Annex III",
  "BXL-ORD-2012_Annex IV",
  "EUR-CON-BER_Annex I",
  "EUR-CON-BER_Annex II",
  "EUR-CON-BER_Annex III",
  "UN-CON-BON-Annex I",
  "UN-CON-BON-Annex II",
  "EUR-DIR-HAB_Annex II",
  "EUR-DIR-HAB_Annex IV"
  # "EUR-DIR-BIR_Annex I"
)
```

Separate them based on annex code and group them in a list of data.frames called `bim_annex`:

```{r list_bim_annexes}
bim_annex <- map(
  bim_annexcode, function(code){
    bim_annex_taxa %>%
      filter(annexcode == code)
    }
)
names(bim_annex) <- annexcode
```

## Read annex data

Import `annexes_gbif_match.csv`:

```{r read_annexes}
annex_df <- read_delim(
  file = here("data", "interim", "annexes_gbif_match.csv"), 
  delim = ",")
```

Keep columns of interest:

```{r select_cols_of_interest}
annex_df <- 
  annex_df %>%
    select(annex_code,
           scientific_name_corrected, 
           scientificName, 
           canonicalName, 
           kingdom, 
           phylum, 
           order, 
           family, 
           genus)
```

Rename columns:
- `scientific_name_corrected` to `annex_scientificName`
- `scientificName` to `gbif_scientificName`
- `canonical` to `gbif_canonicalName`

```{r rename_cols_scientificnames}
annex_df <- 
  annex_df %>%  
  rename("gbif_scientificName" = "scientificName") %>% 
  rename("annex_scientificName" = "scientific_name_corrected") %>%
  rename("gbif_canonicalName" = "canonicalName")
```

Create separate annexes and group them in a list of data.frames called `annex`:

```{r create_list_annex}
annex <- map(annexcode, function(code){
  annex_df %>% filter(annex_code == code)}
)
names(annex) <- annexcode
```

## Read unmatched taxa

Read unmatched taxa in official annexes:

```{r read_unmatched_taxa_annex}
unmatched_annex_taxa_df <- read_csv(
  here("data", "processed", "unmatched_taxa_annexes.csv"),
  na = ""
)
```

Separate them based on annex code and group them in a list of data.frames called `unmatched_annex_taxa`:

```{r make_subsets_unmatched_annex}
unmatched_annex_taxa <- map(
  annexcode, function(code){
    if (code %in% pull(distinct(unmatched_annex_taxa_df, annex_code))) {
      unmatched_annex_taxa_df %>%
      filter(annex_code == code)
    } else {
      tibble(
        annex_code = NA_character_,
        scientific_name_corrected = NA_character_
      )
    }
  }
)
names(unmatched_annex_taxa) <- annexcode
```

# Review BIM annex data

In this section we compare BIM annex data with the information contained in the official annexes.

## Unverifiable taxa 

Some BIM annexes contain taxa which are not verifiable as in the official annex higher ranks are used, such as genus, family and order names. 

### Function to get unverifiable taxa

We write first the function called `get_summarize_unverifiables` to find how many they are. Such function is general and can be aplied to each `bim_*` and the related unverified names and ranks.

```{r define_function_get_summarize_unverifiables}
get_summarize_unverifiables <- function(df, name, rank){
  rank_col <- paste0("gbif_", rank)
  map2_dfr(name,
           rank_col,
           function(name, rank_col) {
             if (!is.na(name)) {
               df %>%
                 filter((!!sym(rank_col)) == name) %>%
                 summarize(number_of_taxa = n()) %>% 
                 mutate(records = name) %>% 
                 select(records, number_of_taxa)
             } else {
               tibble(records = NA_character_,
                      number_of_taxa = NA_integer_)
             }
           }
  )
}
```

### Unverifiable taxa in BXL 2.2

1. Taxa belonging to class `Mammalia` (in annex list: _Alle europese soorten_)
2. Taxa belonging to class `Aves` (in annex list: _Alle europese soorten_)
3. Taxa belonging to class `Amphibia` (in annex list: _All europese soorten_)
4. Taxa belonging to class `Reptilia` (in annex list: _All europese soorten_)
5. Taxa belonging to family `Nymphaeaceae` (in annex list: _Alle Nymphaeaceae soorten waarvan het natuurlijk verspreidingsgebied geheel of gedeeltelijk het nationaal grondgebied omvat_)
6. Taxa belonging to family `Orchidaceae` (in annex list: _Alle Orchidaceae soorten waarvan het natuurlijk verspreidingsgebied geheel of gedeeltelijk het nationaal grondgebied omvatt, met uitsluiting van_ Epipactis helleborine)
7. Taxa belonging to genus `Rosa` (in annex list: _Alle soorten van het geslacht Rosa waarvan het natuurlijk verspreidingsgebied geheel of gedeeltelijk het nationaal grondgebied omvat, met uitzondering van_ Rosa arvensis _en_ Rosa canina L.)

We define them in a data.frame:

```{r unverifiable_taxa_bxl_2_2}
unverifiable_taxa_bxl_2.2 <- tibble(
  rank = c("genus", 
           rep("family", 2),
           rep("class", 4)),
  name = c("Rosa",
           "Orchidaceae",
           "Nymphaeaceae",
           "Mammalia",
           "Aves",
           "Amphibia",
           "Reptilia")
)
```
 
### Unverifiable taxa in BXL 2.3

Taxa belonging to phylum `Bryophyta` (in annex list: _Bryophyta spp._)

```{r unverifiable_taxa_bxl_2.3}
unverifiable_taxa_bxl_2.3 <- tibble(
  rank = "phylum",
  name = "Bryophyta"
)
```

### Unverifiable taxa in BXL 2.5

The following taxa are not included in `annex_bxl_2.5` with their scientific name at species level. So, we are unable to check whether or not their presence in the BIM annex list is correct.

1. Taxa belonging to genus `Coregonus` (in annex list: _Alle europese soorten van Coregonus, behalve Coregonus oxyrhyn_)
2. Taxa belonging to genus `Sphagnum` (in annex list: _Alle europese Sphagnum-soorten_)
3. Taxa belonging to genus `Lycopodium` (in annex list: _Alle europese Lycopodium soorten_)
4. Taxa belonging to genus `Salicornia` ( _Alle Salicornia soorten waarvan het natuurliyk verspreidingsgebied geheel of gedeeltelijk het nationaal grondgebied omvat_)
5. Taxa belonging to family `Acipenseridae` (in annex list: _Alle europese soorten Acipenseridae niet vermeld in bijlage II.2 of II.3_)
6. Taxa belonging to family `macrofunghi` (in annex list: _Alle inheemse macrofunghi soorten_ Epipactis helleborine)
7. Taxa belonging to class `Actinopterygii` (in annex list: _Alle inheemse soorten van vissen niet vermeld in bijlage II.2 of II.3_)
8. _alle inheemse lichenen_ (included in Fungi)

```{r unverifiable_records_bxl_2.5}
unverifiable_taxa_bxl_2.5 <- tibble(
  rank = c(rep("genus", 4),
           "family",
           rep("class", 2),
           "kingdom"),
  name = c("Coregonus",
           "Sphagnum",
           "Lycopodium",
           "Salicornia",
           "Acipenseridae",
           "Elasmobranchii",
           "Actinopterygii",
           "Fungi")
)
```

### Unverifiable taxa in BERN 2

1. Taxa belonging to genus `Hydrobatidae` (in annex list: _all species of Hydrobatidae_)
2. Taxa belonging to genus `Ciconiidae` (in annex list: _all species of Ciconiidae_)
3. Taxa belonging to genus `Otididae` (in annex list: _all species of Otididae_)
4. Taxa belonging to genus `Recurvirostridae` ( _all species of Recurvirostridae_)
5. Taxa belonging to genus `Phalaropodidae` ( _all species of Phalaropodidae_)
6. Taxa belonging to genus `Laridae` ( _all species of Laridae_)
7. Taxa belonging to genus `Pteroclididae` ( _all species of Pteroclididae_)
8. Taxa belonging to family `Threskiornithidae` (in annex list: _all species of Threskiornithidae_)
9. Taxa belonging to family `Gaviidae` (in annex list: _all species of Gaviidae_)
10. Taxa belonging to family `Caprimulgidae` ( _all species of Caprimulgidae_)
11. Taxa belonging to family `Hirundinidae` ( _all species of Hirundinidae_)
12. Taxa belonging to family `Motacillidae` ( _all species of Motacillidae_)
13. Taxa belonging to family `Laniidae` ( _all species of Laniidae_)
14. Taxa belonging to family `Sylviinae` ( _all species of Sylviinae_)
15. Taxa belonging to family `Regulinae` ( _all species of Regulinae_)
16. Taxa belonging to family `Muscicapinae` ( _all species of Muscicapinae_)
17. Taxa belonging to family `Paridae` ( _all species of Paridae_)
18. Taxa belonging to family `Sittidae` ( _all species of Sittidae_)
19. Taxa belonging to family `Certhiidae` ( _all species of Certhiidae_)
20. Taxa belonging to family `Pelicanidae` ( _all species of Pelicanidae_)
21. Taxa belonging to family `Gruidae` ( _all species of Gruidae_)
22. Taxa belonging to order `Falconiformes` ( _all species of Falconiformes_)
23. Taxa belonging to order `Piciformes` ( _all species of Piciformes_)
24. Taxa belonging to order `Strigiformes` ( _all species of Strigiformes_)
25. Taxa belonging to class `Ursidae` (in annex list: _all species of Ursidae_)

We define them in a data.frame:

```{r unverifiable_taxa_bern_2}
unverifiable_taxa_bern_2 <- tibble(
  rank = c(rep("genus", 7), 
    rep("family", 14), 
    rep("order", 3),
    "class"),
  name = c("Hydrobatidae",
    "Ciconiidae",
    "Otididae",
    "Recurvirostridae",
    "Phalaropodidae",
    "Laridae",
    "Pteroclididae",
    "Threskiornithidae",
    "Gaviidae",
    "Caprimulgidae",
    "Hirundinidae",
    "Motacillidae",
    "Laniidae",
    "Sylviinae",
    "Regulinae",
    "Muscicapinae",
    "Paridae",
    "Sittidae",
    "Certhiidae",
    "Pelicanidae",
    "Gruidae",
    "Falconiformes",
    "Piciformes",
    "Strigiformes",
    "Ursidae")
)
```

### Unverifiable taxa in BERN 3

1. Taxa belonging to genus  `Coregonus` (in annex list: _all species of Coregonus _)
2. Taxa belonging to  family `Soricidae` (in annex list: _all species of Soricidae_)
3. Taxa belonging to  family `Gliridae` (in annex list: _all species of Gliridae_)
4. Taxa belonging to  family `Cetacea` (in annex list: _all species of Cetacea not mentioned in Appendix II_)
5. Taxa belonging to family `Cervidae` (in annex list: _all species of Cervidae_)
6. Taxa belonging to  class `Aves` (in annex list: _all species of Birds not included in Appendix II with the exception of Columba palumbus, Corvus corone (corone and/et cornix), Corvus frugilegus, Corvus monedula, Garrulus glandarius, Larus argentatus, Larus fuscus, Larus marinus, Passer domesticus, Sturnus vulgaris, Pica pica_)
7. Taxa belonging to  class `Reptilia` (in annex list: _all species of reptiles not included in Appendix II_)
8. Taxa belonging to  class `Amphibia` (in annex list: _all species of amphibia not included in Appendeix II_)

```{r unverifiable_taxa_bern_3}
unverifiable_taxa_bern_3 <- tibble(
  rank = c("genus",
           rep("family", 4),
           rep("class", 3)),
  name = c("Coregonus",
           "Soricidae",
           "Gliridae",
           "Cetacea",
           "Cervidae",
           "Aves",
           "Reptilia",
           "Amphibia")
)
```

### Unverifiable taxa in BONN 2

All taxa belonging to the following families:

```{r unverifiable_taxa_bonn_2}
unverifiable_taxa_bonn_2 <- tibble(
  rank = rep("family", 68),
  name = c("Acanthizidae",
           "Accipitridae",
           "Acrocephalidae",
           "Aegithalidae",
           "Anatidae",
           "Anseranatidae",
           "Anthropoides",
           "Antigone",
           "Artamidae",
           "Bernieridae",
           "Cathartidae",
           "Cheloniidae",
           "Cinclosomatidae",
           "Cisticolidae",
           "Dasyornithidae",
           "Dermochelyidae",
           "Eulacestomidae",
           "Eupetidae",
           "Falconidae",
           "Falcunculidae",
           "Grus",
           "Haematopodidae",
           "Hyliotidae",
           "Hylocitreidae",
           "Ibidorhynchidae",
           "Ifritidae",
           "Leiotrichidae",
           "Locustellidae",
           "Machaerirhynchidae",
           "Macrosphenidae",
           "Maluridae",
           "Melampittidae",
           "Meliphagidae",
           "Mohouidae",
           "Monarchidae",
           "Motacillidae",
           "Muscicapidae",
           "Oreoicidae",
           "Oriolidae",
           "Orthonychidae",
           "Pachycephalidae",
           "Panuridae",
           "Pellorneidae",
           "Petroicidae",
           "Phoenicopteridae",
           "Phylloscopidae",
           "Picathartidae",
           "Platysteiridae",
           "Pluvianellidae",
           "Pnoepygidae",
           "Polioptilidae",
           "Pomatostomidae",
           "Psophodidae",
           "Recurvirostridae",
           "Regulidae",
           "Rhagologidae",
           "Rhinolophidae",
           "Rhipiduridae",
           "Scolopacidae",
           "Scotocercidae",
           "Stenostiridae",
           "Sylviidae",
           "Timaliidae",
           "Turdidae",
           "Vangidae",
           "Vespertilionidae",
           "Vireonidae",
           "Zosteropidae")
)
```

### Unverifiable taxa in HAB 2

1. Taxa belonging to genus `Eudontomyzon` (in annex list: _Eudontomyzon spp._)
2. Taxa belonging to genus `Alosa` (in annex list: _Alosa spp._)
3. Taxa belonging to genus `Phoxinellus` (in annex list: _Phoxinellus spp._)
4. Taxa belonging to genus `Zingel` (in annex list: _Zingel spp. (except Zingel asper and Zingel zingel)_)

```{r unverifiable_taxa_hab_2}
unverifiable_taxa_hab_2 <- tibble(
  rank = rep("genus", 4),
  name = c("Eudontomyzon",
           "Alosa",
           "Phoxinellus",
           "Zingel")
)
```

### Unverifiable taxa in HAB 4

Taxa belonging to family `Gliridae` (in annex list: _All species except Glis glis and Eliomys quercinus_)

```{r unverifiable_taxa_hab_4}
unverifiable_taxa_hab_4 <- tibble(
  rank = "family",
  name = "Gliridae")
```

### BIM-annexes without unverifiable taxa

The other BIM-annexes don't contain any unverifiable taxa:

```{r unverifiable_taxa}
unverifiable_taxa_empty <- tibble(rank = NA_character_,
                                  name = NA_character_)
unverifiable_taxa_bxl_2.1 <- unverifiable_taxa_empty
unverifiable_taxa_bxl_2.4 <- unverifiable_taxa_empty
unverifiable_taxa_bxl_3 <- unverifiable_taxa_empty
unverifiable_taxa_bxl_4 <- unverifiable_taxa_empty
unverifiable_taxa_bern_1 <- unverifiable_taxa_empty
unverifiable_taxa_bern_2 <- unverifiable_taxa_empty
unverifiable_taxa_bonn_1 <- unverifiable_taxa_empty
```

### Create list of unverifiable taxa

We group all subsets in a list called `unverifiable_taxa`:

```{r list_unverifiable_taxa}
unverifiable_taxa <- map(annexcode, function(code){
  get(paste0("unverifiable_taxa_",code))
})
names(unverifiable_taxa) <- annexcode
```

And we apply function `get_summarize_unverifiables` to all elements of `unverifiable_taxa`. We get a new list of data.frames called `unverifiable_records`, where each of it cointains the number of taxa in the specific BIM-annex for each unverifiable taxon from the correspondent annex:

```{r apply_function_to_all_annexes}
n_taxa_unverifiable_taxa <- map2(bim_annex, unverifiable_taxa,
                             function(df, names) {
                               get_summarize_unverifiables(df, 
                                                           names$name, 
                                                           names$rank)
                             })
names(n_taxa_unverifiable_taxa) <- annexcode
```

Example: unverifiable taxa from BONN 2:

```{r example_unverifiable_bonn_2}
n_taxa_unverifiable_taxa$bonn_2
```

### Total number of unverifiable taxa

For each BIM-annex, we calculate the total number of unverifiable taxa:

```{r calulate_total_unverifiable_records}
n_taxa_unverifiable_taxa_summarized <- map(
  n_taxa_unverifiable_taxa,
  function(df){
    df %>%
      summarize(number_of_taxa = sum(number_of_taxa)) %>%
      mutate(data = "Unverifiable records") %>%
      select(data, number_of_taxa)
  })
names(n_taxa_unverifiable_taxa_summarized) <- annexcode
```

# Create summaries

We define summaries of the match between each pair (BIM-annex, official annex) containing:
1. Number of taxa in `annex_*`
2. Number of taxa in  `annex_*`not matched to GBIF Backbone in `annex_*`
3. Number of taxa in `bim_*`
4. Full match between taxa in both datasets
5. Records present in `bim_*` but not in `annex_*`
6. Unverifiable records in `annex_*`
7. Records present in `annex_*` but not in `bim_*`

## Number of taxa in annexes

Calculate number of taxa in each official annex:

```{r n_taxa_in_annexes}
n_taxa_annex <- map(annex,
                      function(df){
                        df %>% 
                          summarize(number_of_taxa = n()) %>% 
                          mutate(data = "Records annex list") %>% 
                          select(data, number_of_taxa)
                      })
names(n_taxa_annex) <- annexcode
```

## Number of unmatched taxa

Calculate number of unmatched taxa in each annex:

```{r}
n_taxa_unmatched_annex <- map(unmatched_annex_taxa,
                              function(df){
                                df %>% 
                                  summarize(number_of_taxa = n()) %>% 
                                  mutate(data = "GBIF NON matches annex") %>% 
                                  select(data, number_of_taxa)
                              })
names(n_taxa_unmatched_annex) <- annexcode
```

## Number of taxa in BIM-annexes

Calculate number of taxa in each BIM-annex:

```{r n_taxa_in_annex}
n_taxa_bim_annex <- map(bim_annex,
                      function(df){
                        df %>% 
                          summarize(number_of_taxa = n()) %>% 
                          mutate(data = "Records BIM database") %>% 
                          select(data, number_of_taxa)
                      })
names(n_taxa_bim_annex) <- annexcode
```

## Full match

Find number of taxa with full match between each pair (annex, BIM-annex)

```{r n_taxa_full_match}
n_taxa_full_match <- map2(
  annex, bim_annex, function(df1, df2){
    df1 %>% 
      inner_join(df2, by = "gbif_canonicalName") %>% 
      summarize(number_of_taxa = n()) %>% 
      mutate(data = "Full match") %>% 
      select(data, number_of_taxa)
})
names(n_taxa_full_match) <- annexcode
```

## BIM taxa not in annex

Find number of taxa in BIM-annexes not in correspondent annex:

```{r n_taxa_bim_not_in_annex}
n_taxa_bim_not_in_annex <- map2(
  bim_annex, annex, function(df1, df2){
    df1 %>% 
      anti_join(df2, by = c("gbif_canonicalName")) %>%   
      summarize(number_of_taxa = n()) %>% 
      mutate(data = "BIM not in annex") %>% 
      select(data, number_of_taxa)
})
names(n_taxa_bim_not_in_annex) <- annexcode
```

## Annex taxa not in BIM annexes

Find number of taxa in official annexes not in correspondent BIM-annexes:

```{r n_taxa_annex_not_in_bim}
n_taxa_annex_not_in_bim <- map2(
  annex, bim_annex, function(df1, df2){
    df1 %>% 
      anti_join(df2, by = c("gbif_canonicalName")) %>%   
      summarize(number_of_taxa = n()) %>% 
      mutate(data = "Annex not in BIM") %>% 
      select(data, number_of_taxa)
})
names(n_taxa_annex_not_in_bim) <- annexcode
```

## Create summaries

```{r create_summaries}
summaries <- pmap(list(
  n_taxa_annex,
  n_taxa_unmatched_annex,
  n_taxa_bim_annex,
  n_taxa_full_match,
  n_taxa_bim_not_in_annex,
  n_taxa_unverifiable_taxa_summarized,
  n_taxa_annex_not_in_bim),
  bind_rows
)
```

# Ordonnance BXL: Ordonnantie betreffende het natuurbehoud (2012-03-01)

## Annex II.1: Dier- en plantensoorten van communautair belang waarvan de instandhouding de aanwijzing van Natura 2000 gebieden vereist

Summary:

```{r show_summary_2.1}
summaries$bxl_2.1
```

## Annex II.2: Soorten die een strikte bescherming genieten op het hele gewestelijke grondgebied

Summary:

```{r show_summary_2.2}
summaries$bxl_2.2
```

## Annex II.3: Soorten die een geografisch beperkte strikte bescherming genieten

Summary:

```{r show_summary_2.1}
summaries$bxl_2.3
```

## Annex II.4: Soorten van gewestelijk belang

Summary:

```{r show_summary_2.4}
summaries$bxl_2.4
```

## Annex III: Wildsoorten

Summary:

```{r show_summary_bxl_3}
summaries$bxl_3
```

## Annex IV: Invasieve soorten

Summary:

```{r show_summary_bxl_4}
summaries$bxl_4
```

# Bern Convention: Convention on the Conservation of European Wildlife and Natural Habitats 

## Annex I: Strictly protected flora species

Summary:

```{r show_summary_bern_1}
summaries$bern_1
```

## Annex II: Strictly protected fauna species 

Summary:

```{r show_summary_bern_2}
summaries$bern_2
```

## Annex III: Protected fauna species

Summary:

```{r show_summary_bern_3}
summaries$bern_3
```

# BONN Annexes

## Annex I

Summary:

```{r show_summary_bonn_1}
summaries$bonn_1
```

## Annex II

Summary:

```{r show_summary_bonn_2}
summaries$bonn_2
```
