---
title: "How to connect to IBGE - BIM database"
author:
  - Damiano Oldoni
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: yeti
    df_print: paged
knit: (function(input_file, encoding) { rmarkdown::render(input_file, encoding = encoding, output_file = paste0("../docs/",sub(".Rmd", ".html", basename(input_file))))})
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Install software

Before testing connectivity be sure you have installed the following:

1. [PostgreSQL](https://www.postgresql.org/download/)
2. [psqlodbc driver](https://www.postgresql.org/ftp/odbc/versions/msi/) matching the installed PostgreSQL version

You can find general information about accessing PostgreSQL in RStudio [here](https://db.rstudio.com/databases/postgresql/).

## Load libraries

```{r load_libs}
library(odbc)
```

## Check drivers

Check the presence of the PostgreSQL drivers by running:

```{r check-drivers}
drvs <- odbcListDrivers()
```

If you find something like this within data.frame `drvs`, then you ready to connect:

```{r echo=FALSE}
library(stringr)
library(dplyr)
drvs %>%
  filter(str_detect(name, "PostgreSQL Unicode"))
```

# Connection to database

## Store access informations in configuration file

Database is not public, you need to have access rights. A typical way to store database accesses is to make a `config.yml` file in the parent directory "../" of all repositories where access to databases are required. This file would look like this:

```r
default:
  dbname1:
    driver: "my fantastic driver"
    server: "xxx"
    port: 5432
    database: "databasename"
    uid: "username_db1"
    pwd: "password_db1"
  dbname2:
    driver: "PostgreSQL Unicode(x64)"
    server: "xxx"
    port: 1433
    uid: "username_db1"
    pwd = "password_db2"
```

You can retrieve the access infos by `config::get("dbname")`. For example, in order to connect to `dbname1`:  `config::get("dbname1")`. In our case, if you save access informations under name `ibge_bim`, you can retrieve them by:

```r
config::get("ibge_bim")
```

## Connect to database

### Retrieve access information

Retrieve information from configuration file:

```{r get_access_infos}
ibge_bim <- config::get("ibge_bim")
```

### Access to database

We can then connect to database:

```{r connect}
conn <- dbConnect(odbc::odbc(), 
                  driver = ibge_bim$driver,
                  server = ibge_bim$server,
                  database = ibge_bim$database,
                  port = ibge_bim$port,
                  uid = ibge_bim$uid,
                  pwd = ibge_bim$pwd
)
```

### Close connection

Close connection:

```{r close_connection}
dbDisconnect(conn)
```


